---
title: "CTA1 promoter bashing"
author: "JY"
date: "2023-11-24 (updated `r Sys.Date()`)"
output: 
  html_notebook:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r}
library(tidyverse)
library(ggtext)
#library(mgcv)
#library(broom)
library(cowplot)
```

## Goal
Plot the dynamics for CTA1 promoter bashing variants under - Pi and 2 mM H2O2 treatment

## Common plotting functions
**Update 2022-11-08** 
This list is used as a shared plotting set up.
```{r}
p.timecourse <- list(
  stat_summary(fun = "mean", geom = "point", size = 3),
  stat_summary(fun.data = "mean_cl_boot", geom = "errorbar", width = 3),
  stat_smooth(aes(size = Genotype == "WT"), method = "loess", 
              formula = 'y~x', se = FALSE),
  scale_x_continuous(breaks = seq(0,240,30)),
  scale_size_manual(values = c(0.8, 1.5), guide = "none"),
  labs(x = "Time (min)", y = "Cta1-GFP protein level (a.u.)"),
  theme_cowplot(line_size = 0.7, font_size = 14),
  theme(legend.position=c(0.05,0.75), 
        #legend.box.background = element_rect(color = "black"),
        #legend.box.margin = margin(3,3,3,3),
        legend.text = element_text(face = 3), 
        legend.title = element_text(),
        axis.title = element_text(size = rel(1)), 
        #axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1)),
  )
)
```

## CTA1 induction in OL variants

#### -Pi Data
Import data and get the replicate, time and genotype info
```{r}
files <- dir("input/OL_0Pi", pattern = "d01*") # get the file names of the merged
files <- file.path("input//OL_0Pi", files)         # this appends the directory to the file names
names(files) <- c("Rep1-0Pi-Ctrl","Rep2-0Pi-Ctrl", "Rep3-0Pi-Ctrl")

tp <- c("0min",  "30min", "60min", "90min", "120min", "150min",  "180min", "210min", "240min") # ordered time points

# import and process the data to the format we want
data.ol.pi <- map_dfr(files, ~read_csv(., na = c("", "NA", "N/A"), col_types = cols()), .id = "Replicate") %>% 
  #bind_rows(.id = "Replicate") %>% 
  filter(`X Parameter` == "BL1-H", is.na(`Y Parameter`)) %>% 
  select(Replicate, Group, Sample, Count, median = `X Median`, peak = `X Peak`,
         sd = `X SD`, cv = `X %CV`, rCV = `X %rCV`) %>% 
  separate(Group, into = c('Time', 'Treatment'), sep = "-") %>% 
  mutate(Time = ordered(Time, levels = tp)) %>% 
  mutate(Sample = case_when(Sample == "CTA1-GFP-yH298" ~ "yH298-wt",
            Sample == "CTA1-GFP-yH299" ~ "yH299-wt",
            Sample == "yH343-wt" ~ "yH343-rescue",
            Sample == "yH344-wt" ~ "yH344-rescue",
            TRUE ~ Sample)) %>% 
  mutate(Replicate = gsub("\\-.*","", Replicate)) %>% # 
  separate(Sample, into = c("Strain", "genotype"), sep = "-") %>% 
  arrange(Time)
# To replace the value: ifelse() for a single test, case_when() for multiple test
# \\-: escape character "\\" used to escape the special meaning of -, instead take - as a common regex pattern. Or we can specify the - as a regex pattern within a square bracket class [-]
#.*: This part matches any sequence of characters (.* is a common regex pattern for "any character, zero or more times"). metacharacter
```


#### under -Pi
Plot the internal removal variants data(OL)
```{r}
# recode the genotype
levels <- c(WT = "wt", PC = "rescue",`4OL` = "4OL", `6OL` = "6OL", `8OL` = "8OL", `10OL` = "10OL")

pi.dat <- data.ol.pi %>% 
  filter(genotype %in% c("wt", "4OL", "6OL", "8OL", "10OL","rescue")) %>%  
  select(Time, replicate = Replicate, genotype, median) %>%
  mutate(Genotype = fct_recode(genotype, !!!levels)) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  arrange(new_time, genotype, replicate)

genotype_color <- RColorBrewer::brewer.pal(8, "Dark2")[c(8, 7, 6, 1, 5, 4)]  
names(genotype_color) <- names(levels)

pi.dat %>%
  filter(!is.na(new_median)) %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_manual(values= genotype_color, limits = names(genotype_color)) +
  xlab("Phosphate starvation, time (min)")
  #scale_y_continuous(limits = c(NA, 22))

#ggsave("output/OL_0Pi.png", width = 6, height = 8)
```
#### 2mM H2O2 Data
Import 2mM H2O2 OL data and get the replicate, time and genotype info
```{r}
files <- dir("input/OL_2mMH2O2", pattern = "d*") # get the file names of the merged
files <- file.path("input//OL_2mMH2O2", files)      # this appends the directory to the file names
names(files) <- c("Rep1-2mMH2O2","Rep2-2mMH2O2", "Rep3-2mMH2O2", "Rep4-2mMH2O2")

tp <- c("0min",  "30min", "60min", "90min", "120min", "150min",  "180min", "210min", "240min") # ordered time points

# import and process the data to the format we want
data.ol.oxi <- map_dfr(files, ~read_csv(., na = c("", "NA", "N/A"), col_types = cols()), .id = "Replicate") %>% 
  #bind_rows(.id = "Replicate") %>% 
  filter(`X Parameter` == "BL1-H", is.na(`Y Parameter`)) %>% 
  select(Replicate, Group, Sample, Count, median = `X Median`, peak = `X Peak`,
         sd = `X SD`, cv = `X %CV`, rCV = `X %rCV`) %>% 
  separate(Group, into = c('Time', 'Treatment'), sep = "-") %>% 
  mutate(Time = ordered(Time, levels = tp)) %>% 
  mutate(Sample = case_when(Sample == "CTA1-GFP-yH298" ~ "yH298-wt",
            Sample == "CTA1-GFP-yH299" ~ "yH299-wt",
            Sample == "yH343-wt" ~ "yH343-rescue",
            Sample == "yH344-wt" ~ "yH344-rescue",
            TRUE ~ Sample)) %>% 
  mutate(Replicate = gsub("\\-.*","", Replicate)) %>% # 
  separate(Sample, into = c("Strain", "genotype"), sep = "-") %>% 
  arrange(Time)
# To replace the value: ifelse() for a single test, case_when() for multiple test
# \\-: escape character "\\" used to escape the special meaning of -, instead take - as a common regex pattern. Or we can specify the - as a regex pattern within a square bracket class [-]
#.*: This part matches any sequence of characters (.* is a common regex pattern for "any character, zero or more times"). metacharacter
```


#### under 2mM H2O2
Plot the internal removal variants data(OL)
```{r}
# recode the genotype
levels <- c(WT = "wt", PC = "rescue",`4OL` = "4OL", `6OL` = "6OL", `8OL` = "8OL", `10OL` = "10OL")

oxi.ol.dat <- data.ol.oxi %>% 
  filter(genotype %in% c("wt", "4OL", "6OL", "8OL", "10OL","rescue")) %>%  
  select(Time, replicate = Replicate, genotype, median) %>%
  mutate(Genotype = fct_recode(genotype, !!!levels)) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  arrange(new_time, genotype, replicate)

genotype_color <- RColorBrewer::brewer.pal(8, "Dark2")[c(8, 7, 6, 1, 5, 4)]  
names(genotype_color) <- names(levels)

oxi.ol.dat %>%
  filter(!is.na(new_median)) %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_manual(values= genotype_color, limits = names(genotype_color)) +
  xlab("2mM H2O2, time (min)")
  #scale_y_continuous(limits = c(NA, 22))

#ggsave("output/OL_oxi.png", width = 6, height = 8)


##### 2mM H2O2 OL timepoint plot function
tp_GFP <- function(...){
  
  all_tp <- list(...)
  for (input_tp in all_tp) {   
  
    df_tp <- oxi.ol.dat %>% filter(new_time == input_tp) %>% mutate(genotype = factor(genotype, levels = levels))
    
    print(ggplot(df_tp, aes(x = genotype, y = new_median, fill = Genotype)) + 
    geom_point() +
    stat_summary(fun.data = "mean_cl_boot", conf.int = .95, color = "Red", linewidth = 1, size = 0.8) +
    labs(x =paste0("2mM H2O2, OL variants ", deparse1(substitute(input_tp)), " min"), y = "Cta1-GFP protein level (a.u.)") +
    theme_cowplot(line_size = 0.7, font_size = 14) +
    theme(legend.position = "none",
          axis.title = element_text(size = rel(1)), 
          axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1)),))
  }
}

# plot the 2mM H2O2 OL variants at 0min, 90min, 120min timepoints

tp_GFP(0,90,120,240)
  

```



### Import TrunV data and get the replicate, time and genotype info
```{r}
files <- dir("input/TrunV_0Pi", pattern = "d01*") # get the file names of the merged
files <- file.path("input//TrunV_0Pi", files)         # this appends the directory to the file names
names(files) <- c("Rep1-0Pi-Ctrl","Rep2-0Pi-Ctrl", "Rep3-0Pi-Ctrl")

tp <- c("0min",  "30min", "60min", "90min", "120min", "150min",  "180min", "210min", "240min") # ordered time points

# import and process the data to the format we want
data.TrunV <- map_dfr(files, ~read_csv(., na = c("", "NA", "N/A"), col_types = cols()), .id = "Replicate") %>% 
  #bind_rows(.id = "Replicate") %>% 
  filter(!is.na(median),`X Parameter` == "BL1-H", is.na(`Y Parameter`)) %>% 
  select(Replicate, Group, Sample, Count, median = `X Median`, peak = `X Peak`,
         sd = `X SD`, cv = `X %CV`, rCV = `X %rCV`) %>% 
  separate(Group, into = c('Time', 'Treatment'), sep = "-") %>% 
  mutate(Time = ordered(Time, levels = tp)) %>% 
  mutate(Sample = case_when(Sample == "CTA1-GFP-yH298" ~ "yH298-wt",
            Sample == "CTA1-GFP-yH299" ~ "yH299-wt",
            Sample == "yH343-wt" ~ "yH343-rescue",
            Sample == "yH344-wt" ~ "yH344-rescue",
            TRUE ~ Sample)) %>% 
  mutate(Replicate = gsub("\\-.*","", Replicate)) %>%  
  separate(Sample, into = c("Strain", "genotype"), sep = "-") %>% 
  filter(!is.na(median)) %>% 
  arrange(Time) 
# To replace the value: ifelse() for a single test, case_when() for multiple test
# \\-: escape character "\\" used to escape the special meaning of -, instead take - as a common regex pattern. Or we can specify the - as a regex pattern within a square bracket class [-]
#.*: This part matches any sequence of characters (.* is a common regex pattern for "any character, zero or more times"). metacharacter
```



#### Plot the truncation variants data (TrunV)
```{r}
# recode the genotype
levels <- c(WT = "wt", `PC` = "rescue", `2V` = "2V",`4V` = "4V", `6V` = "6V", `8V` = "8V")

pi.dat <- data.TrunV %>% 
  filter(genotype %in% c("wt", "2V", "4V", "6V", "8V","rescue")) %>%  
  select(Time, replicate = Replicate, genotype, median) %>%
  mutate(Genotype = fct_recode(genotype, !!!levels)) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  arrange(new_time, genotype, replicate)

genotype_color <- RColorBrewer::brewer.pal(8, "Dark2")[c(8, 7, 6, 1, 5, 4)]  
names(genotype_color) <- names(levels)

pi.dat %>%
  filter(!is.na(new_median)) %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_manual(values= genotype_color, limits = names(genotype_color)) +
  xlab("Phosphate starvation, time (min)")
  #scale_y_continuous(limits = c(NA, 22))

#ggsave("output/TrunV_0Pi.png", width = 6, height = 8)
```
#### 2mM H2O2 Data
Import 2mM H2O2 TrunV data and get the replicate, time and genotype info
```{r}
files <- dir("input/TrunV_2mMH2O2", pattern = "d*") # get the file names of the merged
files <- file.path("input//TrunV_2mMH2O2", files)      # this appends the directory to the file names
names(files) <- c("Rep1-2mMH2O2","Rep2-2mMH2O2", "Rep3-2mMH2O2")

tp <- c("0min",  "30min", "60min", "90min", "120min", "150min",  "180min", "210min", "240min") # ordered time points

# import and process the data to the format we want
data.TrunV.oxi <- map_dfr(files, ~read_csv(., na = c("", "NA", "N/A"), col_types = cols()), .id = "Replicate") %>% 
  #bind_rows(.id = "Replicate") %>% 
  filter(`X Parameter` == "BL1-H", is.na(`Y Parameter`)) %>% 
  select(Replicate, Group, Sample, Count, median = `X Median`, peak = `X Peak`,
         sd = `X SD`, cv = `X %CV`, rCV = `X %rCV`) %>% 
  separate(Group, into = c('Time', 'Treatment'), sep = "-") %>% 
  mutate(Time = ordered(Time, levels = tp)) %>% 
  mutate(Sample = case_when(Sample == "CTA1-GFP-yH298" ~ "yH298-wt",
            Sample == "CTA1-GFP-yH299" ~ "yH299-wt",
            Sample == "yH343-wt" ~ "yH343-rescue",
            Sample == "yH344-wt" ~ "yH344-rescue",
            TRUE ~ Sample)) %>% 
  mutate(Replicate = gsub("\\-.*","", Replicate)) %>% # 
  separate(Sample, into = c("Strain", "genotype"), sep = "-") %>% 
  filter(!is.na(median)) %>% 
  arrange(Time)
# To replace the value: ifelse() for a single test, case_when() for multiple test
# \\-: escape character "\\" used to escape the special meaning of -, instead take - as a common regex pattern. Or we can specify the - as a regex pattern within a square bracket class [-]
#.*: This part matches any sequence of characters (.* is a common regex pattern for "any character, zero or more times"). metacharacter
```


#### under 2mM H2O2
Plot the Truncation variants data(TrunV)
```{r}
# recode the genotype
levels <- c(WT = "wt", `PC` = "rescue", `2V` = "2V",`4V` = "4V", `6V` = "6V", `8V` = "8V")

oxi.TrunV.dat <- data.TrunV.oxi  %>% 
  filter(genotype %in% c("wt", "2V", "4V", "6V", "8V","rescue")) %>%  
  select(Time, replicate = Replicate, genotype, median) %>%
  mutate(Genotype = fct_recode(genotype, !!!levels)) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  arrange(new_time, genotype, replicate)

genotype_color <- RColorBrewer::brewer.pal(8, "Dark2")[c(8, 7, 6, 1, 5, 4)]  
names(genotype_color) <- names(levels)

oxi.TrunV.dat %>%
  filter(!is.na(new_median)) %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_manual(values= genotype_color, limits = names(genotype_color)) +
  xlab("2mM H2O2 stress, time (min)")
  #scale_y_continuous(limits = c(NA, 22))

#ggsave("output/TrunV_oxi.png", width = 6, height = 8)

##### 2mM H2O2 OL timepoint plot function
tp_GFP <- function(...){
  
  all_tp <- list(...)
  for (input_tp in all_tp) {   
  
    df_tp <- oxi.TrunV.dat %>% filter(new_time == input_tp) %>% mutate(genotype = factor(genotype, levels = levels))
    
    print(ggplot(df_tp, aes(x = genotype, y = new_median, fill = Genotype)) + 
    geom_point() +
    stat_summary(fun.data = "mean_cl_boot", conf.int = .95, color = "Red", linewidth = 1, size = 0.8) +
    labs(x =paste0("2mM H2O2, OL variants ", deparse1(substitute(input_tp)), " min"), y = "Cta1-GFP protein level (a.u.)") +
    theme_cowplot(line_size = 0.7, font_size = 14) +
    theme(legend.position = "none",
          axis.title = element_text(size = rel(1)), 
          axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1)),))
  }
}

# plot the 2mM H2O2 OL variants at 0min, 90min, 120min timepoints

tp_GFP(0,90,120,240)

```
**_1st Differential derivative (Slope) for each curve_**
```{r}
wt.test <- oxi.dat %>% filter(genotype %in% c("wt"))

fit.wt <- loess(new_median~new_time, data=wt.test)

x <- 0:240
px <- predict(fit.wt, newdata = x)
px1 <- diff(px)
#which.max(px1)

wt.loess <- tibble(x = x, y = px)
wt.diff <- tibble (x = x[-1], y = px1)

twoV.test <- oxi.dat %>% filter(genotype %in% c("2V"))

fit.twoV <- loess(new_median~new_time, data=twoV.test)

x <- 0:240
px.2V <- predict(fit.twoV, newdata = x)
px1.2V <- diff(px.2V)

twoV.loess <- tibble(x = x, y = px.2V)
twoV.diff <- tibble (x = x[-1], y = px1.2V)

par(mfrow=c(1, 4))
wt.loess %>% 
ggplot(aes(x = x, y = y)) + geom_point(size = 1, alpha = 0.5) + scale_x_continuous(breaks = seq(0,240,30)) + theme_bw() + ggtitle("loess model")

wt.diff %>% 
  ggplot(aes(x = x, y = y)) + geom_point(size = 1, alpha = 0.5)+ scale_x_continuous(breaks = seq(0, 240, 30)) + theme_bw() + ggtitle("diff(loess model)")


twoV.loess %>% 
ggplot(aes(x = x, y = y)) + geom_point(size = 1, alpha = 0.5) + scale_x_continuous(breaks = seq(0,240,30)) + theme_bw() + ggtitle("loess model")

twoV.diff %>% 
  ggplot(aes(x = x, y = y)) + geom_point(size = 1, alpha = 0.5)+ scale_x_continuous(breaks = seq(0, 240, 30)) + theme_bw() + ggtitle("diff(loess model)")


```
```{r}
fit.wt$


```



**_Statistical test_**

1. Basal level
```{r}
# filter the data and set the basal level
genotype_order <- names(levels)

tmp <- pi.dat %>% 
  filter(new_time == 0, !is.na(new_median)) %>% 
  mutate(Genotype = factor(Genotype, levels = genotype_order))
lm(new_median ~ Genotype, data = tmp) %>% 
  summary()
  #TukeyHSD()
```
2. At 120 minutes
```{r}
tmp <- pi.dat %>% 
  filter(new_time == 120, !is.na(new_median)) %>% 
  mutate(Genotype = factor(Genotype, levels = genotype_order))
lm(new_median ~ Genotype, data = tmp) %>% 
  summary()
```
#### under mock treatment.

Get the subset of Ctrl data
```{r}
ctrl.dat <- data %>% 
  filter(treat == "ctrl", genotype %in% levels) %>%  
  select(Time = Group, Sample, Plate, replicate = Replicate, genotype, median) %>%
  mutate(Genotype = fct_recode(genotype, !!!levels)) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  arrange(new_time, genotype, replicate)
```

```{r}
genotype_color <- RColorBrewer::brewer.pal(8, "Dark2")[c(8, 7, 2)] 
names(genotype_color) <- names(levels)

ctrl.dat %>%
  filter(!is.na(new_median)) %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_manual(values= genotype_color, limits = names(genotype_color)) +
  scale_y_continuous(limits = c(NA, 22))
ggsave("output/20230512-CgSCH9-phosphomimetic-mutant-timecourse-ctrl.png", width = 4.2, height = 4)
```
### Sch9-2E
In _S. cerevisiae_, Pho80/85 was shown to directly phosphorylate Sch9 and thereby regulate its lysosome localization. Here, Jinye interogate a phosphomimetic mutant in _C. glabrata_ and its interaction with _pho80∆_ and _pho81∆_ mutants.
```{r}
levels <- c(WT = "wt", `sch9::Sch9-WT-Nat` = "wt_CgSCH9-Nat", `sch9::Sch9-3E-Nat` = "CgSCH9_3E-Nat", `pho80∆` = "pho80_ko", `pho80∆ sch9::Sch9-2E` = "pho80_ko_Sch9_2E", `pho81∆` = "pho81_ko", `pho81∆ sch9::Sch9-2E` = "pho81_ko_Sch9_2E")

pi.dat.ext <- data %>% 
  filter(treat == "0Pi") %>%  
  select(Time = Group, Sample, Plate, replicate = Replicate, genotype, median) %>%
  mutate(Genotype = fct_recode(genotype,  !!!levels)) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  arrange(new_time, genotype, replicate)
```

Plot the time course
```{r}
genotype_color <- RColorBrewer::brewer.pal(8, "Dark2")[c(8, 7, 6, 1, 5, 4, 2)] 
names(genotype_color) <- names(levels)

pi.dat.ext %>%
  filter(!is.na(new_median)) %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_manual(values= genotype_color, limits = names(genotype_color))
#ggsave("output/20230512-CgSCH9-phosphomimetic-mutant-timecourse-noPi.png", width = 4.2, height = 4)
```

> All comparisons except for rom2- vs WT are statistically significant at 0.05 level after Bonferroni correction (threshold: 0.05/5 = 0.01)

