---
title: "RS_TrunV"
author: "JY"
date: "2024-04-25"
output: html_document
---
```{r}
library(tidyverse)
library(ggtext)
#library(mgcv)
#library(broom)
library(cowplot)
library(ggpattern)
Sys.setenv(LANG = "en")
```

## Goal
Plot the dynamics for CTA1 promoter bashing variants under - Pi and 2 mM H2O2 treatment

## Common plotting functions
**Update 2022-11-08** 
This list is used as a shared plotting set up.
```{r}
p.timecourse <- list(
  stat_summary(fun = "mean", geom = "point", size = 3),
  stat_summary(fun.data = "mean_cl_boot", geom = "errorbar", width = 3),
  stat_smooth(aes(size = Genotype == "WT"), method = "loess", 
              formula = 'y~x', se = FALSE),
  scale_x_continuous(breaks = seq(0,240,30)),
  scale_size_manual(values = c(0.8, 1.5), guide = "none"),
  labs(x = "Time (min)", y = "Cta1-GFP protein level (a.u.)"),
  theme_cowplot(line_size = 0.7, font_size = 14),
  theme(legend.position=c(0.05,0.75), 
        #legend.box.background = element_rect(color = "black"),
        #legend.box.margin = margin(3,3,3,3),
        legend.text = element_text(face = 3), 
        legend.title = element_text(),
        axis.title = element_text(size = rel(1)), 
        #axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1)),
  )
)

mean.timecourse <- list(
  stat_summary(fun = "mean", geom = "point", size = 3),
  stat_smooth(aes(size = Genotype == "WT"), method = "loess", 
              formula = 'y~x', se = FALSE),
  scale_x_continuous(breaks = seq(0,240,30)),
  scale_size_manual(values = c(0.8, 1.5), guide = "none"),
  labs(x = "Time (min)", y = "Cta1-GFP protein level (a.u.)"),
  theme_cowplot(line_size = 0.7, font_size = 14),
  theme(legend.position=c(0.02,0.9),
        #legend.direction = "horizontal",
        #legend.box.background = element_rect(color = "black"),
        #legend.box.margin = margin(3,3,3,3),
        legend.text = element_text(face = 3), 
        legend.title = element_text(),
        axis.title = element_text(size = rel(1)), 
        #axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1)),
  )
)



scatter <- list(stat_summary(fun = "mean", geom = "point", size = 3),
                scale_y_continuous(breaks = seq(0,1.0,0.25)),
                scale_x_continuous(breaks = seq(0,1.0,0.25)),
                scale_size_manual(values = c(0.8, 1.5), guide = "none"),
                expand_limits(x = 0, y = 0),
                theme_cowplot(line_size = 0.7, font_size = 14),
                theme(legend.position=c(0.02,0.82),
                legend.text = element_text(face = 3), 
                legend.title = element_text(),
                axis.title = element_text(size = rel(1)), 
                axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1))))


bar.list <- list(
  stat_summary(fun.data = "mean_cl_boot", conf.int = .95, colour = "red", linewidth = 1, size = 0.8),
  scale_size_manual(values = c(0.8, 1.5), guide = "none"),
  labs(x =paste0("2mM H2O2, OL variants ", deparse1(substitute(input_tp)), " min"), y = "Cta1-GFP protein level (a.u.)"),
  theme_cowplot(line_size = 0.7, font_size = 14),
  theme(legend.position=c(0.02,0.9),
        #legend.direction = "horizontal",
        #legend.box.background = element_rect(color = "black"),
        #legend.box.margin = margin(3,3,3,3),
        legend.text = element_text(face = 3), 
        legend.title = element_text(),
        axis.title = element_text(size = rel(1)), 
        #axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1)),
  )
)

bar.list.no.stat <- list(
  #stat_summary(fun.data = "mean_cl_boot", conf.int = .95, colour = "red", linewidth = 1, size = 0.8),
  scale_size_manual(values = c(0.8, 1.5), guide = "none"),
  theme_cowplot(line_size = 0.7, font_size = 14),
  theme(legend.position=c(0.02,0.9),
        #legend.direction = "horizontal",
        #legend.box.background = element_rect(color = "black"),
        #legend.box.margin = margin(3,3,3,3),
        legend.text = element_text(face = 3), 
        legend.title = element_text(),
        axis.title = element_text(size = rel(1)), 
        #axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1)),
  )
)
  

```



### Prepare data
Import data
Beads were used to adjust for comparable voltage across days
```{r}
files <- dir("input/gz_merge", pattern = "merge-*") # get the file names of the merged
files <- file.path("input/gz_merge", files)         # this appends the directory to the file names

names(files) <- c("Rep1-0Pi-Ctrl","Rep2-0Pi-H2O2","Rep3-0Pi-Ctrl","Rep4-H2O2-Ctrl")


tp <- c("0min",  "30min", "60min", "90min", "120min", "150min",  "180min", "210min", "240min") # ordered time points

# import and process the data to the format we want
dat.ord <- map_dfr(files, ~read_csv(., na = c("", "NA", "N/A"), col_types = cols()), .id = "Replicate") %>% 
  filter(Gate == "CTA1-GFP", `X Parameter` == "BL1-H", is.na(`Y Parameter`), Sample != "beads", Count > 100) %>% 
  select(Replicate, Plate, Group, Sample, Count, median = `X Median`, peak = `X Peak`,
         sd = `X SD`, cv = `X %CV`, rCV = `X %rCV`) %>% 
  mutate(Group = ordered(Group, levels = tp)) %>% 
  arrange(Group)

```

Subset the dataset and merge with the genotype information
```{r}
# get genotype info
tr_files <- dir("input/treatment", pattern = "d0*")
tr_files <- file.path("input/treatment", tr_files)
names(tr_files) <- c("Rep1-0Pi-Ctrl","Rep2-0Pi-H2O2","Rep3-0Pi-Ctrl","Rep4-H2O2-Ctrl")


treat.fil <- tr_files %>% 
  map(~read_csv(., na = c("","NA","N/A"), col_types = cols())) %>% 
  bind_rows(.id = "Replicate") %>% 
  filter(!is.na(`group_name`)) %>% 
  select(Replicate, genotype = `group_name`, Sample, Treatment = treat, HB_strain)


# combine genotype and treat variables into the flow data
data.TrunV.ord <- left_join(dat.ord, treat.fil, by = c("Replicate", "Sample"))
```

#### Import & Combine data
import the RS1 data and combine it with the TrunV data into one dataframe

```{r}
RS1 <- read.csv("input/RS1.csv") %>% mutate(Treatment = treat)

data.comb <- data.TrunV.ord %>% bind_rows(RS1)

# recode the genotype
levels <- c(`WT` = "wt", `PC`= "PC", `IR_2.10` = "2V", `IR_4.10` = "4V", `IR_6.10` = "6V", `IR_8.10` = "8V", `IS_2.4` = "RS1_24",  `IS_2.6` = "RS1_26", `IS_2.8` = "RS1_28", `IS_2.10` = "RS1_210")
G.levels <- c("WT", "PC", "IR_8.10", "IR_6.10", "IR_4.10", "IR_2.10")
W.levels <- c("WT", "PC", "IR_8.10", "IR_6.10", "IR_4.10", "IR_2.10", "IS_2.4", "IS_2.6", "IS_2.8", "IS_2.10")

#levels <- c(`WT` = "wt", `PC` = "PC", `2V` = "2V",`4V` = "4V", `6V` = "6V", `8V` = "8V", `IS_2.4` = "RS1_24",  `IS_2.6` = "RS1_26", `IS_2.8` = "RS1_28", `IS_2.10` = "RS1_210")
```


### separate data based on treatment 
```{r}
pi.dat <- data.comb %>% 
  filter(Treatment == "0Pi") %>%  # , HB_strain != "yH298"
  select(Time = Group, Sample, Plate, replicate = Replicate, genotype, median, HB_strain, Treatment) %>%
  mutate(Genotype = fct_recode(genotype, `WT` = "wt", `PC`= "PC", `IR_2.10` = "2V", `IR_4.10` = "4V", `IR_6.10` = "6V", `IR_8.10` = "8V", `IS_2.4` = "RS1_24",  `IS_2.6` = "RS1_26", `IS_2.8` = "RS1_28", `IS_2.10` = "RS1_210")) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>%
  mutate(Genotype = factor(Genotype, levels = W.levels)) %>% 
  arrange(new_time, genotype, replicate)

h2o2.dat <- data.comb %>% 
  filter(Treatment == "2mMH2O2") %>%  # , HB_strain != "yH298"
  select(Time = Group, Sample, Plate, replicate = Replicate, genotype, median, HB_strain, Treatment) %>%
  mutate(Genotype = fct_recode(genotype, `WT` = "wt", `PC`= "PC", `IR_2.10` = "2V", `IR_4.10` = "4V", `IR_6.10` = "6V", `IR_8.10` = "8V", `IS_2.4` = "RS1_24",  `IS_2.6` = "RS1_26", `IS_2.8` = "RS1_28", `IS_2.10` = "RS1_210")) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  mutate(Genotype = factor(Genotype, levels = W.levels)) %>% 
  arrange(new_time, genotype, replicate)

```

### Get wt, 2V and RS1 CTA1 figure
####induction full time course during -Pi
```{r}
genotype_color <- RColorBrewer::brewer.pal(8, "Dark2")[c(8, 7, 2)] 
#genotype_color <- RColorBrewer::brewer.pal(8, "Dark2")[c(8, 7, 2, 3, 5, 1)] 
names(genotype_color) <- c("WT", "IR_2.10", "IS_2.10")

pi.dat %>%
  filter(!is.na(new_median), Genotype %in% c("WT", "IR_2.10", "IS_2.10")) %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_manual(values= genotype_color, limits = names(genotype_color)) +
  labs(x = "Phosphate starvation, Time (min)")
#ggsave("output/2V_Rs1_0Pi.png", width = 6.3, height = 8)

h2o2.dat %>%
  filter(!is.na(new_median), Genotype %in% c("WT", "IR_2.10", "IS_2.10")) %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_manual(values= genotype_color, limits = names(genotype_color)) +
  labs(x = "2mM H2O2, Time (min)")
#ggsave("output/2V_Rs1_2mMH2O2.png", width = 6.3, height = 8)

```

#### get the reg_infor from TrunV data
#### Get the loess curve fitting for the timecourse data
```{r}
# use the same batch of wt for TrunV
levels.TrunV <- c(`WT`= "wt", `PC` = "PC", `IR_2.10` = "2V", `IR_4.10` = "4V", `IR_6.10` = "6V", `IR_8.10` = "8V")
levels.norescue <- c(`WT` = "wt",`IR_2.10` = "2V", `IR_4.10` = "4V", `IR_6.10` = "6V", `IR_8.10` = "8V")
#genotype_color <- RColorBrewer::brewer.pal(8, "Dark2")[c(8, 7, 6, 1, 5, 4)]  
#names(genotype_color) <- names(levels.TrunV)


pi.TrunV.dat <- data.TrunV.ord %>%
  filter(Treatment == "0Pi") %>%  # , HB_strain != "yH298"
  select(Time = Group, Sample, Plate, Replicate, genotype, median, HB_strain, Treatment) %>%
  mutate(Genotype = fct_recode(genotype, !!!levels)) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  filter(genotype %in% c("wt", "2V", "4V", "6V", "8V","PC")) %>%  
  select(Time, Replicate, genotype, median, Treatment, Strain = HB_strain) %>%
  mutate(Genotype = fct_recode(genotype, !!!levels)) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  mutate(Genotype = factor(Genotype, levels = G.levels)) %>% 
  arrange(new_time, genotype, Replicate)


oxi.TrunV.dat <- data.TrunV.ord %>%
  filter(Treatment == "2mMH2O2") %>%  # , HB_strain != "yH298"
  select(Time = Group, Sample, Plate, Replicate, genotype, median, HB_strain, Treatment) %>%
  mutate(Genotype = fct_recode(genotype, !!!levels)) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  filter(genotype %in% c("wt", "2V", "4V", "6V", "8V","PC")) %>%  
  select(Time, Replicate, genotype, median, Treatment, Strain = HB_strain) %>%
  mutate(Genotype = fct_recode(genotype, !!!levels)) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  mutate(Genotype = factor(Genotype, levels = G.levels)) %>% 
  arrange(new_time, genotype, Replicate)




# original Pi timecourse
pi.TrunV.dat %>%
  filter(!is.na(new_median), genotype != "pc") %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_viridis_d(limits = names(levels.norescue), option = "plasma")  +
  labs(x = "Phosphate starvation, Time (min)")
#ggsave("output/TrunV_0Pi_norescue.png", width = 6, height = 8)

oxi.TrunV.dat %>%
  filter(!is.na(new_median), genotype != "pc") %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_viridis_d(limits = names(levels.norescue), option = "plasma")  +
  xlab("2mM H2O2 stress, time (min)")
  #scale_y_continuous(limits = c(NA, 22))
#ggsave("output/TrunV_oxi_norescue.png", width = 6, height = 8)


# check for the loss fit curve
pi.model <- pi.TrunV.dat  %>% filter(genotype %in% c("wt", "PC", "2V", "4V", "6V", "8V")) %>%  
  group_by(Genotype) %>%
  mutate(model = loess(median ~ new_time, span = 0.75, degree = 2)$fitted)

h2o2.model <- oxi.TrunV.dat  %>% filter(genotype %in% c("wt", "PC", "2V", "4V", "6V", "8V")) %>%  
  group_by(Genotype) %>%
  mutate(model = loess(median ~ new_time, span = 0.75, degree = 2)$fitted)

pi.model %>% ggplot(aes(x = new_time, y = new_median, color = Genotype)) + mean.timecourse + 
  scale_color_viridis_d(limits = names(levels.norescue), option = "plasma")  +
  labs(x = "Phosphate starvation, Time (min)")
#ggsave("output/loess_TrunV_0Pi.png", width = 6.3, height = 8)

h2o2.model %>% ggplot(aes(x = new_time, y = new_median, color = Genotype)) + mean.timecourse + scale_color_viridis_d(limits = names(levels.norescue), option = "plasma")  +
labs(x = "2mM H2O2, Time (min)")
#ggsave("output/loess_TrunV_2mMH2O2.png", width = 6.3, height = 8)
```



```{r}
## get AUC for - Pi data into a dataframe
pi.area <- data.frame(matrix(ncol=2,nrow=0, dimnames=list(NULL, c("pi.area", "Genotype"))))
row_cnt = 1
col_cnt = 1
for (x in names(genotype_color)) {
# calculate AUC  
madeup.pi <- pi.model %>% filter(Genotype == x) %>% loess(formula = new_median ~ new_time, span = 0.75, degree = 2) 
f <- function(x) predict(madeup.pi,newdata=x)
# perform integration
tmp <- integrate(f,0,240)$value
gt_tmp <- x
pi.area[row_cnt,col_cnt] <- tmp
col_cnt = col_cnt + 1
pi.area[row_cnt,col_cnt] <- gt_tmp
col_cnt = 1
row_cnt = row_cnt + 1
}



# get AUC for 2mM H2O2 data
h2o2.area <- data.frame(matrix(ncol=2,nrow=0, dimnames=list(NULL, c("h2o2.area", "Genotype"))))
row_cnt = 1
col_cnt = 1
for (x in names(genotype_color)) {
# calculate AUC  
madeup.h2o2 <- h2o2.model %>% filter(Genotype == x) %>% loess(formula = new_median ~ new_time, span = 0.75, degree = 2) 
f <- function(x) predict(madeup.h2o2,newdata=x)
# perform integration
tmp <- integrate(f,0,240)$value
gt_tmp <- x
h2o2.area[row_cnt,col_cnt] <- tmp
col_cnt = col_cnt + 1
h2o2.area[row_cnt,col_cnt] <- gt_tmp
col_cnt = 1
row_cnt = row_cnt + 1
}

# combind area data
area <- merge(pi.area,h2o2.area)
```

#### Check the loess fitting assumption: normality and curve fitting
```{r}

```






### regulatory information map
```{r}
# calculate the relative information retaining within each genotype
reg_info.TrunV <- area %>% 
  mutate(pi_pcnt = pi.area/(.[.$Genotype == "WT",]$pi.area)) %>% 
  mutate(h2o2_pcnt = h2o2.area/((.[.$Genotype == "WT",]$h2o2.area)))

reg_info.TrunV %>% 
  ggplot(aes(x = pi_pcnt, y = h2o2_pcnt), color = Genotype) +
  scatter + 
  geom_point(size = 3, aes(color = Genotype)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "#666666") +
  labs(x = "Pi (relative to wt)", y = "H2O2 (relative to wt)") +
  scale_color_manual(values = genotype_color, limits = names(genotype_color))
#ggsave("output/TrunVvswt_0pivsh2o2_scatter_Batch2.png", width = 6.3, height = 6.3)
```

# combine TrunV, OL and RS1 reg_info
```{r}
# import OL and RS1 reg_info
# do not need the OL PC, we already have a PC
reg.OL <- read.csv("input/ol_pct.csv") %>% filter(!(Genotype == "PC"))
reg.RS1 <- read.csv("input/rs1_pct.csv")

reg <- reg_info.TrunV %>% bind_rows(reg.OL) %>% bind_rows(reg.RS1)

# annotate contructs with length info
reg_anno <- read.csv("input/reg_anno.csv", sep=",")

# get rid of duplicated WT and do not need PC
reg_all <- left_join(reg, reg_anno, by = "Genotype") %>% select(-c(pi.area, h2o2.area)) %>% unique() %>% filter(!(Genotype == "PC"))


reg_all %>% 
  ggplot(aes(x = pi_pcnt, y = h2o2_pcnt), color = loss_length == 200) +
  scatter + 
  geom_point(aes(color = loss_length == 200 ), size = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "#666666") +
  labs(x = "Pi (relative to wt)", y = "H2O2 (relative to wt)") +
  scale_colour_manual(values = c("#666666", "#ff3399" ))

reg_all %>% 
  ggplot(aes(x = pi_pcnt, y = h2o2_pcnt), color = loss_length) +
  scatter + 
  geom_point(aes(color = loss_length ), size = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "#666666") +
  labs(x = "Pi (relative to wt)", y = "H2O2 (relative to wt)") +
  scale_color_viridis_c()

reg_all %>% 
  ggplot(aes(x = pi_pcnt, y = h2o2_pcnt), color = reg_length) +
  scatter + 
  geom_point(aes(color = reg_length ), size = 3, alpha=0.8) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "#666666") +
  scale_color_viridis_c(option = "plasma") +
  labs(x = expression(AUC[var]/AUC[wt]*", -Pi"), y = expression(AUC[var]/AUC[wt]*",2mM "*H[2]*O[2]))
#ggsave("output/Reg_vswt_scatter.png", width = 6.3, height = 6.3)

reg_all %>% 
  ggplot(aes(x = pi_pcnt, y = h2o2_pcnt), color = type) +
  scatter + 
  geom_point(aes(color = type ), size = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "#666666") +
  labs(x = "Pi (relative to wt)", y = "H2O2 (relative to wt)") +
  scale_color_viridis_d()
```


## Basal expression level of Truncation variants
```{r}
# Use 0min data from both H2O2 and -Pi treatment; 
data.TrunV <- bind_rows(oxi.TrunV.dat, pi.TrunV.dat) %>%  
  select(Time, Replicate, genotype, median, Treatment) %>%
  mutate(Genotype = fct_recode(genotype, !!!levels)) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  arrange(new_time, genotype, Replicate) %>% 
  mutate(genotype = factor(genotype, levels = levels)) %>% 
  mutate(Genotype = factor(Genotype, levels = G.levels))


data.TrunV %>%
  filter(!is.na(new_median), genotype != "PC", Time == "0min") %>% 
  ggplot(aes(x = Genotype, y = new_median, color = Genotype)) + bar.list +
  geom_boxplot() +
  stat_summary(fun.data = "mean_cl_boot", color = "Red", linewidth = 1, size = 0.8) +
  geom_jitter(width = 0.2) +
  scale_color_viridis_d(limits = names(levels.norescue), option = "plasma") +
  theme(legend.position="none") +
  xlab("Truncation Variants, 0 min")
#ggsave("output/TrunV_basal.png", width = 6, height = 8)

## seperately plot the basal level for 0mM Pi and 2mM H2O2
data.TrunV %>%
  filter(!is.na(new_median), genotype != "PC", Time == "0min") %>% 
  ggplot(aes(x = Genotype, y = new_median, color = Genotype)) + bar.list +
  geom_boxplot() +
  stat_summary(fun.data = "mean_cl_boot", color = "Red", linewidth = 1, size = 0.8) +
  geom_jitter(width = 0.2) +
  scale_color_viridis_d(limits = names(levels.norescue), option = "plasma") +
  theme(legend.position=c(0.85,0.9)) +
  xlab("Truncation Variants, 0 min") +
  facet_grid(. ~ Treatment) +
  theme(strip.background = element_rect(colour="black", fill="white", size=1, linetype="solid"), legend.position = "none")
#ggsave("output/TrunV_seperate_basal.png", width = 6, height = 8)

```



## Plateau/maximum expression Level of Truncation variants (V1)
The logic is cancel out the impact of basal expression level of each variants. To do so, minus the 0 min basal expression level from each variants, then plot the deduced data at all tp and the selected tp.
```{r}
# get the individual basal mean
data.TrunV.pi.basal <- pi.TrunV.dat %>% filter(!is.na(median), Time == "0min") %>% group_by(genotype) %>% mutate(basal = median) %>% select(Replicate, Strain, genotype, basal)

data.TrunV.oxi.basal <- oxi.TrunV.dat %>% filter(!is.na(median), Time == "0min") %>% group_by(genotype) %>% mutate(basal = median) %>% select(Replicate, Strain, genotype, basal)


pi.TrunV.dect <- merge(pi.TrunV.dat, data.TrunV.pi.basal, by = c("Strain", "Replicate","genotype")) %>% 
  filter(genotype %in% c("wt", "2V", "4V", "6V", "8V", "rescue"), Strain != "yH298") %>%  
  select(Time, replicate = Replicate, genotype, median, basal) %>%
  mutate(Genotype = fct_recode(genotype, !!!levels)) %>% 
  mutate(new_median =(median-basal)/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  mutate(genotype = factor(genotype, levels = levels)) %>% 
  arrange(new_time, genotype, replicate)

oxi.TrunV.dect <- merge(oxi.TrunV.dat, data.TrunV.oxi.basal, by = c("Strain", "Replicate","genotype")) %>% 
  filter(genotype %in% c("wt", "2V", "4V", "6V", "8V", "rescue"), Strain != "yH298") %>%  
  select(Time, replicate = Replicate, genotype, median, basal) %>%
  mutate(Genotype = fct_recode(genotype, !!!levels)) %>% 
  mutate(new_median =(median-basal)/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  mutate(genotype = factor(genotype, levels = levels)) %>% 
  arrange(new_time, genotype, replicate)

# timecourse deduct curve
pi.TrunV.dect %>%
  filter(!is.na(new_median), genotype != "rescue") %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_viridis_d(limits = names(levels.norescue), option = "plasma")  +
  labs(x = "Phosphate starvation, Time (min)")
#ggsave("output/TrunV_0Pi_deduct.png", width = 6, height = 8)

oxi.TrunV.dect %>%
  filter(!is.na(new_median), genotype != "rescue") %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_viridis_d(limits = names(levels.norescue), option = "plasma")  +
  labs(x = "2mM H2O2, Time (min)")
#ggsave("output/TrunV_oxi_deduct.png", width = 6, height = 8)


# single timepoint deduct curve
# - Pi: 240min
pi.TrunV.dect %>% 
  filter(!is.na(new_median), genotype != "rescue", Time == "240min") %>% 
  ggplot(aes(x = genotype, y = new_median, color = Genotype)) + bar.list +
  geom_boxplot() +
  stat_summary(fun.data = "mean_cl_boot", color = "Red", linewidth = 1, size = 0.8) +
  geom_jitter(width = 0.2) +
  theme(legend.position=c(0.85,0.9)) +
  scale_color_viridis_d(limits = names(levels.norescue), option = "plasma") +
  xlab("Truncation Variants, 0mM Pi, 240 min")
#ggsave("output/TrunV_deduct_0Pi_240.png", width = 6, height = 8)

# 2mM Oxi: 90min
oxi.TrunV.dect %>% 
  filter(!is.na(new_median), genotype != "rescue", Time == "90min") %>% 
  ggplot(aes(x = genotype, y = new_median, color = Genotype)) + bar.list +
  geom_boxplot() +
  stat_summary(fun.data = "mean_cl_boot", color = "Red", linewidth = 1, size = 0.8) +
  geom_jitter(width = 0.2) +
  theme(legend.position=c(0.85,0.9)) +
  scale_color_viridis_d(limits = names(levels.norescue), option = "plasma") +
  xlab("Truncation Variants, 2mM H2O2, 90 min")
#ggsave("output/TrunV_deduct_Oxi_90.png", width = 6, height = 8)

# calculate maximum plateau relative to wt
pi.TrunV.dect.rel <- pi.TrunV.dect %>% filter(Time == "240min") %>% group_by(genotype) %>% 
  mutate(new_mean_ori = mean(median)) %>% 
  mutate(new_mean_red = mean(new_median)) %>% 
  mutate(wt_ori = mean(.[.$Genotype == "WT",]$new_mean_ori)) %>% 
  mutate(wt_red = mean(.[.$Genotype == "WT",]$new_mean_red)) %>% 
  mutate(rel_ori = median/wt_ori) %>% 
  mutate(rel_red = new_median/wt_red) %>% 
  mutate(Treatment = "0Pi")
  
oxi.TrunV.dect.rel <- oxi.TrunV.dect %>% filter(Time == "90min") %>% group_by(genotype) %>% 
  mutate(new_mean_ori = mean(median)) %>% 
  mutate(new_mean_red = mean(new_median)) %>% 
  mutate(wt_ori = mean(.[.$Genotype == "WT",]$new_mean_ori)) %>% 
  mutate(wt_red = mean(.[.$Genotype == "WT",]$new_mean_red)) %>% 
  mutate(rel_ori = median/wt_ori) %>% 
  mutate(rel_red = new_median/wt_red) %>% 
  mutate(Treatment = "2mMH2O2")

TrunV.rel <- bind_rows(pi.TrunV.dect.rel, oxi.TrunV.dect.rel) %>% mutate(Genotype = factor(Genotype, levels = G.levels))

# plot the maximum relative to wt
TrunV.rel %>% 
  filter(!is.na(new_median), genotype != "rescue") %>% 
  ggplot(aes(x = Genotype, y = rel_ori, fill = Treatment, pattern = Treatment)) + bar.list.no.stat +
  geom_bar_pattern(stat = "summary", fun = "mean",
                   position = position_dodge(preserve = "single"),
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,  # 0.1
                   pattern_spacing = 0.04,  # 0.025
                   pattern_key_scale_factor = 1,
                   width=0.8) + 
  scale_fill_manual(values = c("#b4a7d6ff","#f6b26bff")) +
  scale_pattern_manual(values = c(`0Pi` = "stripe", `2mMH2O2` = "none")) +
  geom_errorbar(stat = "summary", fun.data = mean_cl_boot, width = 0.2, color = "Red", linewidth = 1, position = 
                  position_dodge(width = .9)) +
  scale_y_continuous(limits = c(0, 1.5)) +
  #geom_jitter(width = 0.3) +
  geom_hline(yintercept = 1, col = "gray", linetype = "dashed") +
  xlab("Internal removal (IR) variants") +
  ylab("Max induction (relative to wt,ori)") +
  guides(fill = guide_legend(nrow = 1)) +
  theme(strip.background = element_rect(colour="black", fill="white", size=1, linetype="solid"))
#ggsave("output/TrunV_ori_rel_max.png", width = 10.16, height = 2)



TrunV.rel %>% 
  filter(!is.na(new_median), genotype != "rescue") %>% 
  ggplot(aes(x = Genotype, y = rel_red, fill = Treatment, pattern = Treatment)) + bar.list.no.stat +
  geom_bar_pattern(stat = "summary", fun = "mean",
                   position = position_dodge(preserve = "single"),
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,  # 0.1
                   pattern_spacing = 0.04,  # 0.025
                   pattern_key_scale_factor = 1,
                   width=0.8) + 
  scale_fill_manual(values = c("#b4a7d6ff","#f6b26bff")) +
  scale_pattern_manual(values = c(`0Pi` = "stripe", `2mMH2O2` = "none")) +
  geom_errorbar(stat = "summary", fun.data = mean_cl_boot, width = 0.2, color = "Red", linewidth = 1, position = 
                  position_dodge(width = .9)) +
  scale_y_continuous(limits = c(0, 1.5)) +
  #geom_jitter(width = 0.3) +
  geom_hline(yintercept = 1, col = "gray", linetype = "dashed") +
  xlab("Internal removal (IR) variants") +
  ylab("Plateau") +
  guides(fill = guide_legend(nrow = 1)) +
  theme(strip.background = element_rect(colour="black", fill="white", size=1, linetype="solid"), legend.position=c(0.8,0.9))
#ggsave("output/TrunV_red_rel_max.png", width = 10.16, height = 2)

```