---
title: "msn4skn7kos"
author: "JY"
date: "2024-07-16"
output: html_document
---
```{r}
library(tidyverse)
library(ggtext)
#library(mgcv)
library(broom)
library(cowplot)
library(ggpattern)
library(RColorBrewer)
library(tidyverse)
```


## Common plotting functions
**Update 2022-11-08** 
This list is used as a shared plotting set up.
```{r}
p.timecourse <- list(
  stat_summary(fun = "mean", geom = "point", size = 3),
  stat_summary(fun.data = "mean_cl_boot", geom = "errorbar", width = 3),
  stat_smooth(aes(size = Genotype == "WT"), method = "loess", 
              formula = 'y~x', se = FALSE),
  scale_x_continuous(breaks = seq(0,240,30)),
  scale_size_manual(values = c(0.8, 1.5), guide = "none"),
  labs(x = "Time (min)", y = "Cta1-GFP protein level (a.u.)"),
  theme_cowplot(line_size = 0.7, font_size = 14),
  theme(legend.position=c(0.05,0.75), 
        #legend.box.background = element_rect(color = "black"),
        #legend.box.margin = margin(3,3,3,3),
        legend.text = element_text(face = 3), 
        legend.title = element_text(),
        axis.title = element_text(size = rel(1)), 
        #axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1)),
  )
)

bar.list <- list(
  stat_summary(fun.data = "mean_cl_boot", conf.int = .95, colour = "red", linewidth = 1, size = 0.8),
  scale_size_manual(values = c(0.8, 1.5), guide = "none"),
  labs(x =paste0("2mM H2O2, OL variants ", deparse1(substitute(input_tp)), " min"), y = "Cta1-GFP protein level (a.u.)"),
  theme_cowplot(line_size = 0.7, font_size = 14),
  theme(legend.position=c(0.02,0.9),
        #legend.direction = "horizontal",
        #legend.box.background = element_rect(color = "black"),
        #legend.box.margin = margin(3,3,3,3),
        legend.text = element_text(face = 3), 
        legend.title = element_text(),
        axis.title = element_text(size = rel(1)), 
        #axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1)),
  )
)

bar.list.no.stat <- list(
  #stat_summary(fun.data = "mean_cl_boot", conf.int = .95, colour = "red", linewidth = 1, size = 0.8),
  scale_size_manual(values = c(0.8, 1.5), guide = "none"),
  theme_cowplot(line_size = 0.7, font_size = 14),
  theme(legend.position=c(0.02,0.9),
        #legend.direction = "horizontal",
        #legend.box.background = element_rect(color = "black"),
        #legend.box.margin = margin(3,3,3,3),
        legend.text = element_text(face = 3), 
        legend.title = element_text(),
        axis.title = element_text(size = rel(1)), 
        #axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1)),
  )
)
```

## TF deletion mutants compared with wild type
### Prepare data
Import data
```{r}
files <- dir("input/msn24_gz_merge", pattern = "merge-*") # get the file names of the merged
files <- file.path("input/msn24_gz_merge", files)         # this appends the directory to the file names
files.s7y1 <- dir("input/s7y1_gz_merge", pattern = "merge-*") 
files.s7y1 <- file.path("input/s7y1_gz_merge", files.s7y1)  
files.m4s7 <- dir("input/m4s7_gz_merge", pattern = "merge-*") 
files.m4s7 <- file.path("input/m4s7_gz_merge", files.m4s7)


names(files) <- c("Rep1-0Pi-Ctrl","Rep2-0Pi-Rep1-H2O2", "Rep3-0Pi-Rep2-H2O2", "Rep3-H2O2-Ctrl","Rep4-H2O2-Ctrl")
names(files.s7y1) <- c("Rep1-0Pi-Ctrl","Rep2-0Pi-Ctrl","Rep1-H2O2-Rep3-0Pi", "Rep2-H2O2-Rep4-0Pi", "Rep3-H2O2-Ctrl")
names(files.m4s7) <- c("Rep1-0Pi-Ctrl","Rep1-H2O2-Rep2-0Pi", "Rep2-H2O2-Rep3-0Pi", "Rep3-H2O2-Ctrl")


tp <- c("0min",  "30min", "60min", "90min", "120min", "150min",  "180min", "210min", "240min") # ordered time points

# import and process the data to the format we want
dat.ord <- map_dfr(files, ~read_csv(., na = c("", "NA", "N/A"), col_types = cols()), .id = "Replicate") %>% 
  #bind_rows(.id = "Replicate") %>% 
  filter(`X Parameter` == "BL1-H", is.na(`Y Parameter`)) %>% 
  select(Replicate, Plate, Group, Sample, Count, median = `X Median`, peak = `X Peak`,
         sd = `X SD`, cv = `X %CV`, rCV = `X %rCV`) %>% 
  mutate(Group = ordered(Group, levels = tp)) %>% 
  arrange(Group)


# import and process the data to the format we want
dat.ord.s7y1 <- files.s7y1 %>% 
  map(~read_csv(., na = c("", "NA", "N/A"), col_types = cols())) %>% 
  bind_rows(.id = "Replicate") %>% 
  filter(`X Parameter` == "BL1-H", is.na(`Y Parameter`)) %>% 
  select(Replicate, Plate, Group, Sample, Count, median = `X Median`, peak = `X Peak`,
         sd = `X SD`, cv = `X %CV`, rCV = `X %rCV`) %>% 
  mutate(Group = ordered(Group, levels = tp)) %>% 
  arrange(Group)


# import m4s7
dat.ord.m4s7 <- files.m4s7 %>% 
  map(~read_csv(., na = c("", "NA", "N/A"), col_types = cols())) %>% 
  bind_rows(.id = "Replicate") %>% 
  filter(`X Parameter` == "BL1-H", is.na(`Y Parameter`)) %>% 
  select(Replicate, Plate, Group, Sample, Count, median = `X Median`, peak = `X Peak`,
         sd = `X SD`, cv = `X %CV`, rCV = `X %rCV`) %>% 
  mutate(Group = ordered(Group, levels = tp)) %>% 
  arrange(Group)
```

Subset the dataset and merge with the genotype information
```{r}
# get genotype info
tr_files <- dir("input/msn24_treatment", pattern = "d0*")
tr_files <- file.path("input/msn24_treatment", tr_files)
names(tr_files) <- c("Rep1-0Pi-Ctrl","Rep2-0Pi-Rep1-H2O2", "Rep3-0Pi-Rep2-H2O2", "Rep3-H2O2-Ctrl","Rep4-H2O2-Ctrl")

tr_files.s7y1 <- dir("input/s7y1_treatment", pattern = "d0*")
tr_files.s7y1 <- file.path("input/s7y1_treatment", tr_files.s7y1)
names(tr_files.s7y1) <- c("Rep1-0Pi-Ctrl","Rep2-0Pi-Ctrl","Rep1-H2O2-Rep3-0Pi", "Rep2-H2O2-Rep4-0Pi", "Rep3-H2O2-Ctrl")

tr_files.m4s7 <- dir("input/m4s7_treatment", pattern = "d0*")
tr_files.m4s7 <- file.path("input/m4s7_treatment", tr_files.m4s7)
names(tr_files.m4s7) <- c("Rep1-0Pi-Ctrl","Rep1-H2O2-Rep2-0Pi", "Rep2-H2O2-Rep3-0Pi", "Rep3-H2O2-Ctrl")



treat.fil <- tr_files %>% 
  map(~read_csv(., na = c("","NA","N/A"), col_types = cols())) %>% 
  bind_rows(.id = "Replicate") %>% 
  filter(!is.na(`group_name`)) %>% 
  select(Replicate, genotype = `group_name`, Sample, Strain, treat)


treat.fil.s7y1 <- tr_files.s7y1 %>% 
  map(~read_csv(., na = c("","NA","N/A"), col_types = cols())) %>% 
  bind_rows(.id = "Replicate") %>% 
  filter(!is.na(`group_name`)) %>% 
  select(Replicate, genotype = `group_name`, Sample, Strain, treat)


treat.fil.m4s7 <- tr_files.m4s7 %>% 
  map(~read_csv(., na = c("","NA","N/A"), col_types = cols())) %>% 
  bind_rows(.id = "Replicate") %>% 
  filter(!is.na(`group_name`)) %>% 
  select(Replicate, genotype = `group_name`, Sample, Strain, treat)




# combine genotype and treat variables into the flow data
data <- left_join(dat.ord, treat.fil, by = c("Replicate", "Sample"))
data.s7y1 <- left_join(dat.ord.s7y1, treat.fil.s7y1, by = c("Replicate", "Sample"))
data.m4s7 <- left_join(dat.ord.m4s7, treat.fil.m4s7, by = c("Replicate", "Sample"))


data.tmp <- add_row(data,data.s7y1)
data.cb <- add_row(data.tmp, data.m4s7)


```

### CTA1 induction during -Pi [Fig. 4A]
Get the subset of no Pi data
```{r}
pi.dat <- data.cb %>% 
  filter(treat == "0Pi" & !genotype %in% c("skn7yap1_ko","msn2msn4_ko")) %>%  
  select(Time = Group, Sample, Plate, replicate = Replicate, genotype, Strain, median) %>%
  mutate(Genotype = fct_recode(genotype, WT = "wt", `msn2-` = "msn2_ko", `msn4-` = "msn4_ko", `yap1-` = "yap1_ko",
                               `skn7-` = "skn7_ko", `msn4- skn7-` = "msn4skn7_ko")) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  arrange(new_time, genotype, replicate)
```

Get the subset of 2mM H2O2 data
```{r}
h2o2.dat <- data.cb %>% 
  filter(treat == "H2O2" & !genotype %in% c("skn7yap1_ko","msn2msn4_ko")) %>%  
  select(Time = Group, Sample, Plate, replicate = Replicate, genotype, Strain, median) %>%
  mutate(Genotype = fct_recode(genotype, WT = "wt", `msn2-` = "msn2_ko", `msn4-` = "msn4_ko", `yap1-` = "yap1_ko",
                               `skn7-` = "skn7_ko", `msn4- skn7-` = "msn4skn7_ko")) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  arrange(new_time, genotype, replicate)
```
*need to check the skn7 and yap1 data, I got them switched somehow*

#### Full time course

```{r}
genotype_color <- RColorBrewer::brewer.pal(8, "Dark2")[c(8, 7, 2, 3, 5, 1)] 
names(genotype_color) <- c("WT", "msn2-", "yap1-", "msn4-","skn7-", "msn4- skn7-")

pi.dat %>%
  filter(!is.na(new_median)) %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_manual(values= genotype_color, limits = names(genotype_color))
#ggsave("output/20221108-trans-deletion-timecourse-noPi.png", width = 4.2, height = 4)
```

### Basal vs relative induction levels [Fig. S?]
From the graph above, we can see that the tf∆ mutants affected both the basal expression level (at 0 min) and the fold change (relative to the 0'). To further dissect the effects on these two components, we will make some additional plots below
```{r}
genotype_order <- c("WT", "msn2-", "yap1-", "msn4-", "skn7-", "msn4- skn7-")
genotype_color <- RColorBrewer::brewer.pal(8, "Dark2")[c(8, 7, 2, 3, 5, 1)] 
names(genotype_color) <- genotype_order
p.base <- pi.dat %>% 
  filter(new_time == 0, !is.na(new_median)) %>% 
  mutate(Genotype = factor(Genotype, levels = genotype_order)) %>% 
  ggplot(aes(x = Genotype, y = new_median)) +
  geom_point(position = position_jitter(width = 0.1), alpha = 0.5, shape = 16) +
  stat_summary(aes(color = Genotype), fun.data = "mean_cl_boot", geom = "errorbar",
               width = 0.3, linewidth = 1.2) +
  stat_summary(aes(color = Genotype), fun = "mean", geom = "point", size = 4) +
  scale_color_manual(values = genotype_color, guide = "none") +
  labs(y = "Cta1-GFP protein level (a.u.)") +
  theme_cowplot(line_size = 0.7, font_size = 14) +
  background_grid() + panel_border(color = "black", size = 1.5) +
  theme(#legend.position=c(0.05,0.75), 
        #legend.box.background = element_rect(color = "black"),
        #legend.box.margin = margin(3,3,3,3),
        #legend.text = element_text(face = 3), 
        #legend.title = element_text(),
    axis.line = element_blank(),
    axis.title = element_text(size = rel(1)), 
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), axis.text = element_text(size = rel(1)),
  )

p.base
```

Statistical test
```{r}
# filter the data and set the basal level
tmp <- pi.dat %>% 
  filter(new_time == 0, !is.na(new_median)) %>% 
  mutate(Genotype = factor(Genotype, levels = genotype_order))
lm(new_median ~ Genotype, data = tmp) %>% 
  summary()
  #TukeyHSD()
```
> All comparisons except for msn2- vs WT are statistically significant at 0.05 level after Bonferroni correction (threshold: 0.05/5 = 0.01)

### Max Slope
#### plot & calculate the slope/differiation changes of msn4 and skn7:- Pi
```{r}
# function to plot the max slope from the raw data frame
max_slope <- function(genotypes, rawdata) {
  # Ensure genotypes is a vector
  if (!is.vector(genotypes)) {
    stop("genotypes should be a vector of genotype names.")
  }
  
  # Initiate an empty tibble dataframe to store the results
  result <- tibble(
    Genotype = character(),
    Strain = character(),
    Max_Slope_tp = numeric(),
    Max_Slope = numeric()
  )

  # Iterate through each genotype
  for (geno in genotypes) {
    # Filter data for the current genotype
    filtered_data <- rawdata %>%
      filter(Genotype == geno)
    
    if (nrow(filtered_data) == 0) {
      cat("No data found for genotype:", geno, "\n")
      next
    }
    
    for (z in unique(filtered_data$Strain)) {
    # Fit the loess model
    strain_data <- filtered_data %>% filter(Strain == z)   
    test.pi <- loess(new_median ~ new_time, data =  strain_data, span = 0.75, degree = 2)
    
    # Create a new data frame for prediction
    newdata <- data.frame(new_time = 0:240)
    px <- predict(test.pi, newdata = newdata)
    
    # Handle potential NA values in prediction
    if (any(is.na(px))) {
      px <- zoo::na.approx(px)
    }
    
    px1 <- diff(px)
    
    max_slope_index <- which.max(px1)
    max_slope_value <- max(px1, na.rm = TRUE)
    

    # Append results to the tibble
    result <- result %>% add_row(
      Genotype = geno,
      Strain = z,
      Max_Slope_tp = max_slope_index,
      Max_Slope = max_slope_value
    )
    
    # Plotting
    par(mfrow = c(1, 2))
    
    plot(0:240, px, type = "l", main = paste("loess model for", geno), xlab = "Time", ylab = "Fitted values")
    plot(1:240, px1, type = "l", main = paste("diff(loess model) for", geno), xlab = "Time", ylab = "Slope")
    abline(v = max_slope_index, col = "red")
  }}
  
  return(result)
}

# Example usage
max_slope(c("WT","msn4-", "skn7-", "msn4- skn7-"), pi.dat) # view the plot
pi.slope <- max_slope(c("WT","msn4-", "skn7-", "msn4- skn7-"), pi.dat) # save the tibble output to a new variable


G.levels <- c("WT","msn4-", "skn7-", "msn4- skn7-")

# get the maximum slope relative to the wt
pi.slope.rel <- pi.slope %>% filter(!is.na(Max_Slope)) %>% 
  group_by(Genotype) %>% 
  mutate(mean_slope = mean(Max_Slope)) %>%
  mutate(wt_slope = mean(.[.$Genotype == "WT",]$mean_slope)) %>% 
  mutate(rel_slope = Max_Slope/wt_slope) %>% 
  mutate(Treatment = "0Pi") %>% 
  mutate(Genotype = factor(Genotype, levels = G.levels)) %>% 
  select(-wt_slope)

# plot the maximum slope relative to wt
pi.slope.rel %>% 
  filter(!is.na(rel_slope)) %>% 
  ggplot(aes(x = Genotype, y = rel_slope, fill = Treatment, pattern = Treatment)) + bar.list.no.stat +
 geom_bar_pattern(stat = "summary", fun = "mean",
                   position = position_dodge(preserve = "single"),
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,  # 0.1
                   pattern_spacing = 0.04,  # 0.025
                   pattern_key_scale_factor = 1,
                   width=0.8) + 
  scale_fill_manual(values = c("#b4a7d6ff")) +
  scale_pattern_manual(values = c(`0Pi` = "stripe")) +
  geom_errorbar(stat = "summary", fun.data = mean_cl_boot, width = 0.2, color = "Black", linewidth = 1, position = 
                  position_dodge(width = .9)) +
  scale_y_continuous(limits = c(0, 1.5)) +
  #geom_jitter(width = 0.3) +
  geom_hline(yintercept = 1, col = "gray", linetype = "dashed") +
  xlab("Internal removal (IR) variants") +
  ylab("Max Slope") +
  guides(fill = guide_legend(nrow = 1)) +
  theme(strip.background = element_rect(colour="black", fill="white", size=1, linetype="solid"), legend.position= "none")
#ggsave("output/msn4skn7_slope_0Pi.png", width = 10.16, height = 2)

```
### Epistasis test between msn4 and Skn7
#### Max slope
```{r}
slope_matrix <- pi.slope.rel %>% mutate(beta_Msn4 = 1, beta_Skn7 = 1)

slope_matrix$beta_Msn4 <- with(slope_matrix, ifelse(Genotype == "msn4-" | Genotype == "msn4- skn7-", 0, 1))
slope_matrix$beta_Skn7 <- with(slope_matrix, ifelse(Genotype == "skn7-" | Genotype == "msn4- skn7-", 0, 1))

fit.pi <- lm(rel_slope ~  beta_Msn4 + beta_Skn7 + beta_Msn4:beta_Skn7, data = slope_matrix)
round(summary(fit.pi)$coef, 6)

# only main effect model
fit.main_effect <- lm(rel_slope ~  beta_Msn4 + beta_Skn7, data = slope_matrix)
round(summary(fit.main_effect)$coef, 6)

# Conclusion: adding the interaction term does not affecting the intepretation for the main effect (no interaction term dominating the model)

# Extract the summary of the fit
fit_summary <- summary(fit.pi)

# Create a tibble to store the coefficients and their statistics
coefficients_tb <- tibble(
  main_effect = rownames(fit_summary$coefficients),
  Estimate = fit_summary$coefficients[, "Estimate"],
  SE = fit_summary$coefficients[, "Std. Error"],
  t_value = fit_summary$coefficients[, "t value"],
  p_value = fit_summary$coefficients[, "Pr(>|t|)"]
)



fct.levels <- c("b0","msn4", "skn7", "msn4:skn7")
# Round the coefficients for better readability (optional)
coefficients_tb <- coefficients_tb %>% mutate(across(where(is.numeric), ~round(.x, 6))) %>% 
                   mutate(feature = "slope", Treatment = "0Pi") %>% 
                   mutate(seg = fct_recode(main_effect, 'msn4' = "beta_Msn4", 'skn7' = "beta_Skn7", 'msn4:skn7' = "beta_Msn4:beta_Skn7", 
                                           'b0' = "(Intercept)")) %>% 
                   mutate(seg = factor(seg, levels = fct.levels))


coefficients_tb %>% filter(main_effect != "(Intercept)") %>% 
  ggplot(aes(x = seg, y = Estimate, fill = Treatment, pattern = Treatment)) + bar.list.no.stat +
  geom_bar_pattern(stat = "summary", fun = "mean",
                   position = position_dodge(preserve = "single"),
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,  # 0.1
                   pattern_spacing = 0.04,  # 0.025
                   pattern_key_scale_factor = 1,
                   width=0.8) + 
  scale_fill_manual(values = c("#b4a7d6ff"),labels = c("-Pi")) +
  scale_pattern_manual(values = c(`0Pi` = "stripe"),labels = c("-Pi")) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.2, color = "Black", linewidth = 1, 
                position = position_dodge(width = 0.9)) +
  scale_y_continuous(breaks = seq(-0.4, 1.0, 0.2)) +
  coord_cartesian(ylim = c(-0.2, 1)) +
  geom_hline(yintercept = 1, col = "gray", linetype = "dashed") +
  xlab("promoter bashing variants") +
  ylab("Max Slope") +
  guides(fill = guide_legend(nrow = 1, title = NULL), pattern = guide_legend(title = NULL)) +
  theme(strip.background = element_rect(colour="black", fill="white", size=1, linetype="solid"), 
        legend.position= c(0.02, 0.9),
        axis.title.x=element_blank())
ggsave("output/msn4skn7_max_slope_0Pi.png", width = 3.3, height = 3.3)
```


### Basal
```{r}
# all genotype level
all.levels <- c("WT", "msn2-", "msn4-", "skn7-", "msn4- skn7-", "yap1-" )

## plot the 0mM Pi and 2mM H2O2 basal side by side
dat <- data.cb %>% filter(!genotype %in% c("skn7yap1_ko","msn2msn4_ko")) %>%  
  select(Time = Group, Sample, Plate, replicate = Replicate, genotype, Strain, median, treat) %>%
  mutate(Genotype = fct_recode(genotype, WT = "wt", `msn2-` = "msn2_ko", `msn4-` = "msn4_ko", `yap1-` = "yap1_ko",
                               `skn7-` = "skn7_ko", `msn4- skn7-` = "msn4skn7_ko")) %>%
  mutate(Genotype = factor(Genotype, levels = all.levels)) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  arrange(new_time, genotype, replicate)

# raw basal
dat %>% 
  filter(!is.na(new_median), Genotype %in% c("WT", "msn4-", "skn7-", "msn4- skn7-"), Time == "0min", treat != "ctrl") %>% 
  ggplot(aes(x = Genotype, y = new_median, fill = treat, pattern = treat)) + bar.list.no.stat +
  geom_bar_pattern(stat = "summary", fun = "mean",
                   position = position_dodge(preserve = "single"),
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,  # 0.1
                   pattern_spacing = 0.04,  # 0.025
                   pattern_key_scale_factor = 1,
                   width=0.8) + 
  scale_fill_manual(values = c("#b4a7d6ff","#f6b26bff")) +
  scale_pattern_manual(values = c(`0Pi` = "stripe", `H2O2` = "none")) +
  geom_errorbar(stat = "summary", fun.data = mean_cl_boot, width = 0.2, color = "Black", linewidth = 1, position = 
                  position_dodge(width = .9)) +
  scale_y_continuous(limits = c(0, 10)) +
  #geom_jitter(width = 0.3) +
  geom_hline(yintercept = 6, col = "gray", linetype = "dashed") +
  xlab("Internal removal (IR) variants") +
  ylab("Basal exp (a.u.)") +
  guides(fill = guide_legend(nrow = 1)) +
  theme(strip.background = element_rect(colour="black", fill="white", size=1, linetype="solid"), legend.position = "none")
#ggsave("output/basal_bar.png", width = 10.16, height = 2)

# calculate relative basal expression changes to wt
bas.rel.pi <- dat %>%
          filter(!is.na(new_median), Genotype %in% c("WT", "msn4-", "skn7-", "msn4- skn7-"), Time == "0min", treat == "0Pi") %>% 
          group_by(genotype) %>% 
          mutate(new_mean_base = mean(median)) %>% 
          mutate(wt_basal = mean(.[.$Genotype == "WT",]$new_mean_base)) %>% 
          mutate(rel_base = median/wt_basal) %>% 
          mutate(Treatment = "0Pi") %>% 
          select(-wt_basal)


bas.rel.oxi <- dat %>% 
          filter(!is.na(new_median), Genotype %in% c("WT", "msn4-", "skn7-", "msn4- skn7-"), Time == "0min", treat == "H2O2") %>% 
          group_by(genotype) %>% 
          mutate(new_mean_base = mean(median)) %>% 
          mutate(wt_basal = mean(.[.$Genotype == "WT",]$new_mean_base)) %>% 
          mutate(rel_base = median/wt_basal) %>% 
          mutate(Treatment = "2mMH2O2") %>% 
          select(-wt_basal)

bas.rel <- bind_rows(bas.rel.oxi, bas.rel.pi)
bas.rel.save <- bas.rel %>% select(Time, genotype, Genotype, rel_base)
```
## Plateau/maximum expression
The logic is cancel out the impact of basal expression level of each variants. To do so, minus the 0 min basal expression level from each variants, then plot the deduced data at all tp and the selected tp.
```{r}
# get the individual basal mean
data.pi.basal <- pi.dat %>% filter(!is.na(median), Time == "0min") %>% group_by(genotype) %>% mutate(basal = median) %>% select(replicate, Strain, genotype, basal)

data.oxi.basal <- h2o2.dat %>% filter(!is.na(median), Time == "0min") %>% group_by(genotype) %>% mutate(basal = median) %>% select(replicate, Strain, genotype, basal)


# Define levels and G.levels as vectors
levels <- c("wt", "msn4_ko", "skn7_ko", "msn4skn7_ko")
G.levels <- c("WT", "msn4-", "skn7-", "msn4- skn7-")


# calculate the new CTA1-GFP (a.u.) by decuding the 0 min number for individual strain
pi.dect <- merge(pi.dat, data.pi.basal, by = c("Strain", "replicate","genotype")) %>% 
  filter(Genotype %in% c("WT", "msn4-", "skn7-", "msn4- skn7-")) %>%  
  select(Time, replicate, genotype, median, basal, Genotype) %>%
  mutate(new_median =(median-basal)/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  mutate(genotype = factor(genotype, levels = levels)) %>% 
  mutate(Genotype = factor(Genotype, levels = G.levels)) %>% 
  arrange(new_time, genotype, replicate)

oxi.dect <- merge(h2o2.dat, data.oxi.basal, by = c("Strain", "replicate","genotype")) %>% 
  filter(Genotype %in% c("WT", "msn4-", "skn7-", "msn4- skn7-")) %>%  
  select(Time, replicate, genotype, median, basal, Genotype) %>%
  mutate(new_median =(median-basal)/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  mutate(genotype = factor(genotype, levels = levels)) %>% 
  mutate(Genotype = factor(Genotype, levels = G.levels)) %>% 
  arrange(new_time, genotype, replicate)


# timecourse deduct curve
pi.dect %>%
  filter(!is.na(new_median)) %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_viridis_d(limits = names(G.levels))  +
  labs(x = "Phosphate starvation, Time (min)")
#ggsave("output/msn4skn7_0Pi_deduct.png", width = 6, height = 8)

oxi.dect %>%
  filter(!is.na(new_median)) %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_viridis_d(limits = names(G.levels))  +
  labs(x = "2mM H2O2, Time (min)")
#ggsave("output/msn4skn7_oxi_deduct.png", width = 6, height = 8)


# raw single timepoint deduct curve
# - Pi: 240min
pi.dect %>% 
  filter(!is.na(new_median), Time == "240min") %>% 
  ggplot(aes(x = Genotype, y = new_median, color = Genotype)) + bar.list +
  geom_boxplot() +
  stat_summary(fun.data = "mean_cl_boot", color = "Red", linewidth = 1, size = 0.8) +
  geom_jitter(width = 0.2) +
  scale_color_viridis_d(limits = names(G.levels)) +
  theme(legend.position = "none") +
  xlab("0mM Pi, 240 min")
#ggsave("output/OL_deduct_0Pi_240.png", width = 6, height = 8)


# calculate maximum plateau relative to wt: - Pi
pi.dect.rel <- pi.dect %>% filter(Time == "240min") %>% group_by(genotype) %>% 
  mutate(new_mean_ori = mean(median)) %>% 
  mutate(new_mean_red = mean(new_median)) %>% 
  mutate(wt_ori = mean(.[.$Genotype == "WT",]$new_mean_ori)) %>% 
  mutate(wt_red = mean(.[.$Genotype == "WT",]$new_mean_red)) %>% 
  mutate(rel_ori = median/wt_ori) %>% 
  mutate(rel_red = new_median/wt_red) %>% 
  mutate(Treatment = "0Pi") %>% 
  select(-c(wt_ori, wt_red, rel_ori))



# plot the basal deduced maximum plateau relative to wt
pi.dect.rel %>% 
  filter(!is.na(new_median)) %>% 
  ggplot(aes(x = Genotype, y = rel_red, fill = Treatment, pattern = Treatment)) + bar.list.no.stat +
  geom_bar_pattern(stat = "summary", fun = "mean",
                   position = position_dodge(preserve = "single"),
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,  # 0.1
                   pattern_spacing = 0.04,  # 0.025
                   pattern_key_scale_factor = 1,
                   width=0.8) + 
  scale_fill_manual(values = c("#b4a7d6ff","#f6b26bff"),labels = c("-Pi", bquote(~H[2]*O[2]))) +
  scale_pattern_manual(values = c(`0Pi` = "stripe", `2mMH2O2` = "none"),labels = c("-Pi", bquote(~H[2]*O[2]))) +
  geom_errorbar(stat = "summary", fun.data = mean_cl_boot, width = 0.2, color = "Black", linewidth = 1, position = 
                  position_dodge(width = .9)) +
  scale_y_continuous(limits = c(0, 1.5)) +
  #geom_jitter(width = 0.3) +
  geom_hline(yintercept = 1, col = "gray", linetype = "dashed") +
  xlab("Internal removal (IR) variants") +
  ylab("Plateau") +
  guides(fill = guide_legend(nrow = 1, title = NULL), pattern = guide_legend(title = NULL)) +
  theme(strip.background = element_rect(colour="black", fill="white", size=1, linetype="solid"), legend.position=c(0.8,0.9))
#ggsave("output/msn4skn7_red_rel_max.png", width = 10.16, height = 2)

```


### Epistasis test between msn4 and Skn7
#### Deduced Plateau
```{r}
plateau_matrix <- pi.dect.rel %>% mutate(beta_Msn4 = 1, beta_Skn7 = 1)

plateau_matrix$beta_Msn4 <- with(plateau_matrix, ifelse(Genotype == "msn4-" | Genotype == "msn4- skn7-", 0, 1))
plateau_matrix$beta_Skn7 <- with(plateau_matrix, ifelse(Genotype == "skn7-" | Genotype == "msn4- skn7-", 0, 1))

fit.pi <- lm(rel_red ~  beta_Msn4 + beta_Skn7 + beta_Msn4:beta_Skn7, data = plateau_matrix)
round(summary(fit.pi)$coef, 6)


# only main effect model
fit.main_effect <- lm(rel_red ~  beta_Msn4 + beta_Skn7, data = plateau_matrix)
round(summary(fit.main_effect)$coef, 6)

# Conclusion: adding the interaction term reduces the 50% main effect size of Skn7 and 30% of Msn4...

# check multicollinearity:
library(car)
vif(fit.pi)

# Centering and Scalling to reduce multicollinearity
plateau_matrix$beta_Msn4_centered <- scale(plateau_matrix$beta_Msn4, center = TRUE, scale = TRUE)
plateau_matrix$beta_Skn7_centered <- scale(plateau_matrix$beta_Skn7, center = TRUE, scale = TRUE)

# Fitting the model with centered variables
fit.pi_centered <- lm(rel_red ~ beta_Msn4_centered * beta_Skn7_centered, data = plateau_matrix)
summary(fit.pi_centered)

# Checking for multicollinearity again (all main effect's vif are smaller and close to one)
vif(fit.pi_centered)


# Extract the summary of the fit
fit_summary <- summary(fit.pi)

# Create a tibble to store the coefficients and their statistics
coefficients_tb <- tibble(
  main_effect = rownames(fit_summary$coefficients),
  Estimate = fit_summary$coefficients[, "Estimate"],
  SE = fit_summary$coefficients[, "Std. Error"],
  t_value = fit_summary$coefficients[, "t value"],
  p_value = fit_summary$coefficients[, "Pr(>|t|)"]
)


fct.levels <- c("b0","msn4", "skn7", "msn4:skn7")
# Round the coefficients for better readability (optional)
coefficients_tb <- coefficients_tb %>% mutate(across(where(is.numeric), ~round(.x, 6))) %>% 
                   mutate(feature = "deduced_Plateau", Treatment = "0Pi") %>% 
                   mutate(seg = fct_recode(main_effect, 'msn4' = "beta_Msn4", 'skn7' = "beta_Skn7", 'msn4:skn7' = "beta_Msn4:beta_Skn7", 
                                           'b0' = "(Intercept)")) %>% 
                   mutate(seg = factor(seg, levels = fct.levels))


coefficients_tb %>% filter(main_effect != "(Intercept)") %>% 
  ggplot(aes(x = seg, y = Estimate, fill = Treatment, pattern = Treatment)) + bar.list.no.stat +
  geom_bar_pattern(stat = "summary", fun = "mean",
                   position = position_dodge(preserve = "single"),
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,  # 0.1
                   pattern_spacing = 0.04,  # 0.025
                   pattern_key_scale_factor = 1,
                   width=0.8) + 
  scale_fill_manual(values = c("#b4a7d6ff"),labels = c("-Pi")) +
  scale_pattern_manual(values = c(`0Pi` = "stripe"),labels = c("-Pi")) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.2, color = "Black", linewidth = 1, 
                position = position_dodge(width = 0.9)) +
  scale_y_continuous(breaks = seq(-0.4, 1.0, 0.2)) +
  coord_cartesian(ylim = c(-0.2, 1)) +
  geom_hline(yintercept = 1, col = "gray", linetype = "dashed") +
  xlab("promoter bashing variants") +
  ylab("Plateau") +
  guides(fill = guide_legend(nrow = 1, title = NULL), pattern = guide_legend(title = NULL)) +
  theme(strip.background = element_rect(colour="black", fill="white", size=1, linetype="solid"), 
        legend.position= "none",
        axis.title.x=element_blank())
#ggsave("output/msn4skn7_ded_plateau_0Pi.png", width = 3.3, height = 3.3)
```



### Basal
```{r}
# combine - Pi and H2O2 basal
# calculate maximum plateau relative to wt: - Pi
pi.bas.rel <- pi.dect %>% filter(Time == "0min", !is.na(new_median)) %>% group_by(genotype) %>% 
               mutate(new_mean_base = mean(median)) %>% 
               mutate(wt_basal = mean(.[.$Genotype == "WT",]$new_mean_base)) %>% 
               mutate(rel_base = median/wt_basal) %>% 
               mutate(Treatment = "0Pi") #%>% 
               #select(-c(wt_basal, new_mean_base))

oxi.bas.rel <- oxi.dect %>% filter(Time == "0min", !is.na(new_median)) %>% group_by(genotype) %>% 
               mutate(new_mean_base = mean(median)) %>% 
               mutate(wt_basal = mean(.[.$Genotype == "WT",]$new_mean_base)) %>% 
               mutate(rel_base = median/wt_basal) %>% 
               mutate(Treatment = "2mMH2O2") 

bas.rel <- bind_rows(pi.bas.rel, oxi.bas.rel)

# raw base
bas.rel %>% 
  filter(!is.na(rel_base), Time == "0min") %>% 
  ggplot(aes(x = Genotype, y = rel_base, fill = Treatment, pattern = Treatment)) + bar.list.no.stat +
  geom_bar_pattern(stat = "summary", fun = "mean",
                   position = position_dodge(preserve = "single"),
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,  # 0.1
                   pattern_spacing = 0.04,  # 0.025
                   pattern_key_scale_factor = 1,
                   width=0.8) + 
  scale_fill_manual(values = c("#b4a7d6ff","#f6b26bff")) +
  scale_pattern_manual(values = c(`0Pi` = "stripe", `2mMH2O2` = "none")) +
  geom_errorbar(stat = "summary", fun.data = mean_cl_boot, width = 0.2, color = "Black", linewidth = 1, position = 
                  position_dodge(width = .9)) +
  scale_y_continuous(limits = c(0, 1.5)) +
  #geom_jitter(width = 0.3) +
  geom_hline(yintercept = 1, col = "gray", linetype = "dashed") +
  xlab("Internal removal (IR) variants") +
  ylab("Basal exp") +
  guides(fill = guide_legend(nrow = 1)) +
  theme(strip.background = element_rect(colour="black", fill="white", size=1, linetype="solid"), legend.position = "none")
#ggsave("output/Ir_basal_rel_bar.png", width = 10.16, height = 2)

## linear modeling
bas_matrix <- bas.rel %>% mutate(beta_Msn4 = 1, beta_Skn7 = 1)

bas_matrix$beta_Msn4 <- with(bas_matrix, ifelse(Genotype == "msn4-" | Genotype == "msn4- skn7-", 0, 1))
bas_matrix$beta_Skn7 <- with(bas_matrix, ifelse(Genotype == "skn7-" | Genotype == "msn4- skn7-", 0, 1))

# 0 Pi
bas_matrix.pi <- bas_matrix %>% filter(Treatment == "0Pi")
fit.pi <- lm(rel_base ~  beta_Msn4 + beta_Skn7 + beta_Msn4:beta_Skn7, data = bas_matrix.pi)
round(summary(fit.pi)$coef, 6)

# check multicollinearity:
library(car)
vif(fit.pi)

# Centering and Scalling to reduce multicollinearity
bas_matrix.pi$beta_Msn4_centered <- scale(bas_matrix.pi$beta_Msn4, center = TRUE, scale = TRUE)
bas_matrix.pi$beta_Skn7_centered <- scale(bas_matrix.pi$beta_Skn7, center = TRUE, scale = TRUE)

# Fitting the model with centered variables
fit.pi_centered <- lm(rel_base ~ beta_Msn4_centered * beta_Skn7_centered, data = bas_matrix.pi)
summary(fit.pi_centered)

# Checking for multicollinearity again (all main effect's vif are smaller and close to one, the conclusion does not change)
vif(fit.pi_centered)


# Extract the summary of the fit
fit_summary <- summary(fit.pi)

# Create a tibble to store the coefficients and their statistics
coefficients_tb_pi <- tibble(
  main_effect = rownames(fit_summary$coefficients),
  Estimate = fit_summary$coefficients[, "Estimate"],
  SE = fit_summary$coefficients[, "Std. Error"],
  t_value = fit_summary$coefficients[, "t value"],
  p_value = fit_summary$coefficients[, "Pr(>|t|)"]
)


# H2O2
bas_matrix.h2o2 <- bas_matrix %>% filter(Treatment == "2mMH2O2")
fit.h2o2 <- lm(rel_base ~  beta_Msn4 + beta_Skn7 + beta_Msn4:beta_Skn7, data = bas_matrix.h2o2)
round(summary(fit.h2o2)$coef, 6)

fit_summary <- summary(fit.h2o2)

# Create a tibble to store the coefficients and their statistics
coefficients_tb_h2o2 <- tibble(
  main_effect = rownames(fit_summary$coefficients),
  Estimate = fit_summary$coefficients[, "Estimate"],
  SE = fit_summary$coefficients[, "Std. Error"],
  t_value = fit_summary$coefficients[, "t value"],
  p_value = fit_summary$coefficients[, "Pr(>|t|)"]
)


fct.levels <- c("b0","msn4", "skn7", "msn4:skn7")
# Round the coefficients for better readability (optional)
coefficients_tb_pi <- coefficients_tb_pi %>% mutate(across(where(is.numeric), ~round(.x, 6))) %>% 
                   mutate(feature = "deduced_Plateau", Treatment = "0Pi") %>% 
                   mutate(seg = fct_recode(main_effect, 'msn4' = "beta_Msn4", 'skn7' = "beta_Skn7", 'msn4:skn7' = "beta_Msn4:beta_Skn7", 
                                           'b0' = "(Intercept)")) %>% 
                   mutate(seg = factor(seg, levels = fct.levels))

coefficients_tb_h2o2 <- coefficients_tb_h2o2 %>% mutate(across(where(is.numeric), ~round(.x, 6))) %>% 
                   mutate(feature = "deduced_Plateau", Treatment = "2mMH2O2") %>% 
                   mutate(seg = fct_recode(main_effect, 'msn4' = "beta_Msn4", 'skn7' = "beta_Skn7", 'msn4:skn7' = "beta_Msn4:beta_Skn7", 
                                           'b0' = "(Intercept)")) %>% 
                   mutate(seg = factor(seg, levels = fct.levels))


coefficients_bas <- bind_rows(coefficients_tb_pi, coefficients_tb_h2o2)

coefficients_bas %>% filter(main_effect != "(Intercept)") %>% 
  ggplot(aes(x = seg, y = Estimate, fill = Treatment, pattern = Treatment)) + bar.list.no.stat +
  geom_bar_pattern(stat = "summary", fun = "mean",
                   position = position_dodge(preserve = "single"),
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,  # 0.1
                   pattern_spacing = 0.04,  # 0.025
                   pattern_key_scale_factor = 1,
                   width=0.8) + 
 scale_fill_manual(values = c("#b4a7d6ff","#f6b26bff"),labels = c("-Pi", bquote(~H[2]*O[2]))) +
  scale_pattern_manual(values = c(`0Pi` = "stripe", `2mMH2O2` = "none"),labels = c("-Pi", bquote(~H[2]*O[2]))) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.2, color = "Black", linewidth = 1, 
                position = position_dodge(width = 0.9)) +
  scale_y_continuous(breaks = seq(-0.4, 1.0, 0.2)) +
  coord_cartesian(ylim = c(-0.2, 1)) +
  geom_hline(yintercept = 1, col = "gray", linetype = "dashed") +
  xlab("TF knockout") +
  ylab("Basal Exp") +
  guides(fill = guide_legend(nrow = 1, title = NULL), pattern = guide_legend(title = NULL)) +
  theme(strip.background = element_rect(colour="black", fill="white", size=1, linetype="solid"), 
        legend.position= c(0.02, 0.9),
        axis.title.x=element_blank())
ggsave("output/msn4skn7_base_rel.png", width = 3.3, height = 3.3)


```

# Expression epistasis analysis on Yap1 and Skn7 during Oxidative stress
Goal: positive control for my linear model
```{r}
h2o2.dat.all <- data.cb %>% 
  filter(treat == "H2O2") %>%  
  select(Time = Group, Sample, Plate, replicate = Replicate, genotype, Strain, median) %>%
  mutate(Genotype = fct_recode(genotype, WT = "wt", `msn2-` = "msn2_ko", `msn4-` = "msn4_ko", `yap1-` = "yap1_ko",
                               `skn7-` = "skn7_ko", `msn4- skn7-` = "msn4skn7_ko", `msn2- msn4-` = "msn2msn4_ko", `yap1- skn7-` = "skn7yap1_ko")) %>% 
  mutate(new_median =median/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  arrange(new_time, genotype, replicate)


genotype_color <- RColorBrewer::brewer.pal(8, "Dark2")[c(8, 7, 2, 3, 5, 1, 4, 6)] 
names(genotype_color) <- c("WT", "msn2-", "yap1-", "msn4-","skn7-", "msn4- skn7-", "yap1- skn7-", "msn2- msn4-")

h2o2.dat.all %>%
  filter(!is.na(new_median)) %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_manual(values= genotype_color, limits = c("WT", "msn2-", "msn4-", "msn2- msn4-",
                                "yap1- skn7-", "msn4- skn7-", "yap1-", "skn7-"))
#ggsave("output/20221108-trans-deletion-timecourse-h2o2.png", width = 4.2, height = 4)

```
# max slope
### Max Slope
#### plot & calculate the slope/differiation changes of msn4 and skn7:- Pi
```{r}
# function to plot the max slope from the raw data frame
max_slope <- function(genotypes, rawdata) {
  # Ensure genotypes is a vector
  if (!is.vector(genotypes)) {
    stop("genotypes should be a vector of genotype names.")
  }
  
  # Initiate an empty tibble dataframe to store the results
  result <- tibble(
    Genotype = character(),
    Strain = character(),
    Max_Slope_tp = numeric(),
    Max_Slope = numeric()
  )

  # Iterate through each genotype
  for (geno in genotypes) {
    # Filter data for the current genotype
    filtered_data <- rawdata %>%
      filter(Genotype == geno)
    
    if (nrow(filtered_data) == 0) {
      cat("No data found for genotype:", geno, "\n")
      next
    }
    
    for (z in unique(filtered_data$Strain)) {
    # Fit the loess model
    strain_data <- filtered_data %>% filter(Strain == z)   
    test.pi <- loess(new_median ~ new_time, data =  strain_data, span = 0.75, degree = 2)
    
    # Create a new data frame for prediction
    newdata <- data.frame(new_time = 0:240)
    px <- predict(test.pi, newdata = newdata)
    
    # Handle potential NA values in prediction
    if (any(is.na(px))) {
      px <- zoo::na.approx(px)
    }
    
    px1 <- diff(px)
    
    max_slope_index <- which.max(px1)
    max_slope_value <- max(px1, na.rm = TRUE)
    

    # Append results to the tibble
    result <- result %>% add_row(
      Genotype = geno,
      Strain = z,
      Max_Slope_tp = max_slope_index,
      Max_Slope = max_slope_value
    )
    
    # Plotting
    par(mfrow = c(1, 2))
    
    plot(0:240, px, type = "l", main = paste("loess model for", geno), xlab = "Time", ylab = "Fitted values")
    plot(1:240, px1, type = "l", main = paste("diff(loess model) for", geno), xlab = "Time", ylab = "Slope")
    abline(v = max_slope_index, col = "red")
  }}
  
  return(result)
}

# Example usage
max_slope(c("WT","yap1-", "skn7-", "yap1- skn7-"), h2o2.dat.all) # view the plot
h2o2.slope <- max_slope(c("WT","yap1-", "skn7-", "yap1- skn7-"), h2o2.dat.all) # save the tibble output to a new variable


G.levels <- c("WT","yap1-", "skn7-", "yap1- skn7-")

# get the maximum slope relative to the wt
h2o2.slope.rel <- h2o2.slope %>% filter(!is.na(Max_Slope)) %>% 
  group_by(Genotype) %>% 
  mutate(mean_slope = mean(Max_Slope)) %>%
  mutate(wt_slope = mean(.[.$Genotype == "WT",]$mean_slope)) %>% 
  mutate(rel_slope = Max_Slope/wt_slope) %>% 
  mutate(Treatment = "2mMH2O2") %>% 
  mutate(Genotype = factor(Genotype, levels = G.levels)) %>% 
  select(-wt_slope)

# plot the maximum slope relative to wt
h2o2.slope.rel %>% 
  filter(!is.na(rel_slope)) %>% 
  ggplot(aes(x = Genotype, y = rel_slope, fill = Treatment, pattern = Treatment)) + bar.list.no.stat +
 geom_bar_pattern(stat = "summary", fun = "mean",
                   position = position_dodge(preserve = "single"),
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,  # 0.1
                   pattern_spacing = 0.04,  # 0.025
                   pattern_key_scale_factor = 1,
                   width=0.8) + 
 scale_fill_manual(values = c("#b4a7d6ff","#f6b26bff"),labels = c("-Pi", bquote(~H[2]*O[2]))) +
  scale_pattern_manual(values = c(`0Pi` = "stripe", `2mMH2O2` = "none"),labels = c("-Pi", bquote(~H[2]*O[2]))) +
  geom_errorbar(stat = "summary", fun.data = mean_cl_boot, width = 0.2, color = "Black", linewidth = 1, position = 
                  position_dodge(width = .9)) +
  scale_y_continuous(limits = c(0, 1.5)) +
  #geom_jitter(width = 0.3) +
  geom_hline(yintercept = 1, col = "gray", linetype = "dashed") +
  xlab("Internal removal (IR) variants") +
  ylab("Max Slope") +
  guides(fill = guide_legend(nrow = 1)) +
  theme(strip.background = element_rect(colour="black", fill="white", size=1, linetype="solid"), legend.position= "none")
#ggsave("output/skn7yap1_slope_h2o2.png", width = 10.16, height = 2)

```
### Epistasis test between msn4 and Skn7
#### Max slope
```{r}
slope_matrix <- h2o2.slope.rel %>% mutate(beta_Skn7 = 1, beta_Yap1 = 1)

slope_matrix$beta_Yap1 <- with(slope_matrix, ifelse(Genotype == "yap1-" | Genotype == "yap1- skn7-", 0, 1))
slope_matrix$beta_Skn7 <- with(slope_matrix, ifelse(Genotype == "skn7-" | Genotype == "yap1- skn7-", 0, 1))

fit.h2o2 <- lm(rel_slope ~  beta_Skn7 + beta_Yap1 + beta_Skn7:beta_Yap1, data = slope_matrix)
round(summary(fit.h2o2)$coef, 6)

# Extract the summary of the fit
fit_summary <- summary(fit.h2o2)

# Create a tibble to store the coefficients and their statistics
coefficients_tb <- tibble(
  main_effect = rownames(fit_summary$coefficients),
  Estimate = fit_summary$coefficients[, "Estimate"],
  SE = fit_summary$coefficients[, "Std. Error"],
  t_value = fit_summary$coefficients[, "t value"],
  p_value = fit_summary$coefficients[, "Pr(>|t|)"]
)



fct.levels <- c("b0","yap1", "skn7", "yap1:skn7")
# Round the coefficients for better readability (optional)
coefficients_tb <- coefficients_tb %>% mutate(across(where(is.numeric), ~round(.x, 6))) %>% 
                   mutate(feature = "slope", Treatment = "0Pi") %>% 
                   mutate(seg = fct_recode(main_effect, 'yap1' = "beta_Yap1", 'skn7' = "beta_Skn7", 'yap1:skn7' = "beta_Skn7:beta_Yap1", 
                                           'b0' = "(Intercept)")) %>% 
                   mutate(seg = factor(seg, levels = fct.levels))


coefficients_tb %>% filter(main_effect != "(Intercept)") %>% 
  ggplot(aes(x = seg, y = Estimate, fill = Treatment, pattern = Treatment)) + bar.list.no.stat +
  geom_bar_pattern(stat = "summary", fun = "mean",
                   position = position_dodge(preserve = "single"),
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,  # 0.1
                   pattern_spacing = 0.04,  # 0.025
                   pattern_key_scale_factor = 1,
                   width=0.8) + 
  scale_fill_manual(values = c("#b4a7d6ff"),labels = c("-Pi")) +
  scale_pattern_manual(values = c(`0Pi` = "stripe"),labels = c("-Pi")) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.2, color = "Black", linewidth = 1, 
                position = position_dodge(width = 0.9)) +
  scale_y_continuous(breaks = seq(-0.4, 1.0, 0.2)) +
  coord_cartesian(ylim = c(-0.2, 1)) +
  geom_hline(yintercept = 1, col = "gray", linetype = "dashed") +
  xlab("promoter bashing variants") +
  ylab("Max Slope") +
  guides(fill = guide_legend(nrow = 1, title = NULL), pattern = guide_legend(title = NULL)) +
  theme(strip.background = element_rect(colour="black", fill="white", size=1, linetype="solid"), 
        legend.position= c(0.02, 0.9),
        axis.title.x=element_blank())
#ggsave("output/skn7yap1_max_slope_0Pi.png", width = 3.3, height = 3.3)
```
## Plateau/maximum expression
The logic is cancel out the impact of basal expression level of each variants. To do so, minus the 0 min basal expression level from each variants, then plot the deduced data at all tp and the selected tp.
```{r}
# get the individual basal mean

data.oxi.basal <- h2o2.dat.all %>% filter(!is.na(median), Time == "0min") %>% group_by(genotype) %>% mutate(basal = median) %>% select(replicate, Strain, genotype, basal)


# Define levels and G.levels as vectors
levels <- c("wt", "yap1_ko", "skn7_ko", "skn7yap1_ko")
G.levels <- c("WT", "yap1-", "skn7-", "yap1- skn7-")


# calculate the new CTA1-GFP (a.u.) by decuding the 0 min number for individual strain
oxi.dect <- merge(h2o2.dat.all, data.oxi.basal, by = c("Strain", "replicate","genotype")) %>% 
  filter(Genotype %in% c("WT", "yap1-", "skn7-", "yap1- skn7-")) %>%  
  select(Time, replicate, genotype, median, basal, Genotype) %>%
  mutate(new_median =(median-basal)/1000) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time)))%>% 
  mutate(genotype = factor(genotype, levels = levels)) %>% 
  mutate(Genotype = factor(Genotype, levels = G.levels)) %>% 
  arrange(new_time, genotype, replicate)


# timecourse deduct curve

oxi.dect %>%
  filter(!is.na(new_median)) %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_viridis_d(limits = names(G.levels))  +
  labs(x = "2mM H2O2, Time (min)")
#ggsave("output/msn4skn7_oxi_deduct.png", width = 6, height = 8)


# raw single timepoint deduct curve
# - Pi: 240min
oxi.dect %>% 
  filter(!is.na(new_median), Time == "90min") %>% 
  ggplot(aes(x = Genotype, y = new_median, color = Genotype)) + bar.list +
  geom_boxplot() +
  stat_summary(fun.data = "mean_cl_boot", color = "Red", linewidth = 1, size = 0.8) +
  geom_jitter(width = 0.2) +
  scale_color_viridis_d(limits = names(G.levels)) +
  theme(legend.position = "none") +
  xlab("0mM Pi, 240 min")
#ggsave("output/OL_deduct_0Pi_240.png", width = 6, height = 8)


# calculate maximum plateau relative to wt: - Pi
oxi.dect.rel <- oxi.dect %>% filter(Time == "90min") %>% group_by(genotype) %>% 
  mutate(new_mean_ori = mean(median)) %>% 
  mutate(new_mean_red = mean(new_median)) %>% 
  mutate(wt_ori = mean(.[.$Genotype == "WT",]$new_mean_ori)) %>% 
  mutate(wt_red = mean(.[.$Genotype == "WT",]$new_mean_red)) %>% 
  mutate(rel_ori = median/wt_ori) %>% 
  mutate(rel_red = new_median/wt_red) %>% 
  mutate(Treatment = "0Pi") %>% 
  select(-c(wt_ori, wt_red, rel_ori))



# plot the basal deduced maximum plateau relative to wt
oxi.dect.rel %>% 
  filter(!is.na(new_median)) %>% 
  ggplot(aes(x = Genotype, y = rel_red, fill = Treatment, pattern = Treatment)) + bar.list.no.stat +
  geom_bar_pattern(stat = "summary", fun = "mean",
                   position = position_dodge(preserve = "single"),
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,  # 0.1
                   pattern_spacing = 0.04,  # 0.025
                   pattern_key_scale_factor = 1,
                   width=0.8) + 
  scale_fill_manual(values = c("#b4a7d6ff","#f6b26bff"),labels = c("-Pi", bquote(~H[2]*O[2]))) +
  scale_pattern_manual(values = c(`0Pi` = "stripe", `2mMH2O2` = "none"),labels = c("-Pi", bquote(~H[2]*O[2]))) +
  geom_errorbar(stat = "summary", fun.data = mean_cl_boot, width = 0.2, color = "Black", linewidth = 1, position = 
                  position_dodge(width = .9)) +
  scale_y_continuous(limits = c(0, 1.5)) +
  #geom_jitter(width = 0.3) +
  geom_hline(yintercept = 1, col = "gray", linetype = "dashed") +
  xlab("Internal removal (IR) variants") +
  ylab("Plateau") +
  guides(fill = guide_legend(nrow = 1, title = NULL), pattern = guide_legend(title = NULL)) +
  theme(strip.background = element_rect(colour="black", fill="white", size=1, linetype="solid"), legend.position=c(0.8,0.9))
#ggsave("output/msn4skn7_red_rel_max.png", width = 10.16, height = 2)

```


### Epistasis test between msn4 and Skn7
#### Deduced Plateau
```{r}
plateau_matrix <- oxi.dect.rel %>% mutate(beta_Yap1 = 1, beta_Skn7 = 1)

plateau_matrix$beta_Yap1 <- with(plateau_matrix, ifelse(Genotype == "yap1-" | Genotype == "yap1- skn7-", 0, 1))
plateau_matrix$beta_Skn7 <- with(plateau_matrix, ifelse(Genotype == "skn7-" | Genotype == "yap1- skn7-", 0, 1))


# variable selection does not solve the main effect dominating the linear model problem


# Method set 2: alternative method (take out the interaction term)
fit.main_effect <- lm(rel_red ~  beta_Yap1 + beta_Skn7, data = plateau_matrix)
round(summary(fit.main_effect)$coef, 6)


fit.oxi <- lm(rel_red ~  beta_Yap1 + beta_Skn7 + beta_Yap1:beta_Skn7, data = plateau_matrix)
round(summary(fit.oxi)$coef, 6)


# check multicollinearity:
library(car)
vif(fit.oxi) # interaction terms has vif >6 suggesting high multicollinearity

# Centering and Scalling to reduce multicollinearity
plateau_matrix$beta_Yap1_centered <- scale(plateau_matrix$beta_Yap1, center = TRUE, scale = TRUE)
plateau_matrix$beta_Skn7_centered <- scale(plateau_matrix$beta_Skn7, center = TRUE, scale = TRUE)

# Fitting the model with centered variables
fit.oxi_centered <- lm(rel_red ~ beta_Yap1_centered * beta_Skn7_centered, data = plateau_matrix)
summary(fit.oxi_centered)

# Checking for multicollinearity again (this time the main effect is significant, vif close to 1, the centering&scalling changed the intepretation of the model with interaction terms)
vif(fit.oxi_centered)



# Extract the summary of the fit
fit_summary <- summary(fit.oxi)

# Create a tibble to store the coefficients and their statistics
coefficients_tb <- tibble(
  main_effect = rownames(fit_summary$coefficients),
  Estimate = fit_summary$coefficients[, "Estimate"],
  SE = fit_summary$coefficients[, "Std. Error"],
  t_value = fit_summary$coefficients[, "t value"],
  p_value = fit_summary$coefficients[, "Pr(>|t|)"]
)


fct.levels <- c("b0","yap1", "skn7", "yap1:skn7")
# Round the coefficients for better readability (optional)
coefficients_tb <- coefficients_tb %>% mutate(across(where(is.numeric), ~round(.x, 6))) %>% 
                   mutate(feature = "deduced_Plateau", Treatment = "2mMH2O2") %>% 
                   mutate(seg = fct_recode(main_effect, 'yap1' = "beta_Yap1", 'skn7' = "beta_Skn7", 'yap1:skn7' = "beta_Yap1:beta_Skn7", 
                                           'b0' = "(Intercept)")) %>% 
                   mutate(seg = factor(seg, levels = fct.levels))


coefficients_tb %>% filter(main_effect != "(Intercept)") %>% 
  ggplot(aes(x = seg, y = Estimate, fill = Treatment, pattern = Treatment)) + bar.list.no.stat +
  geom_bar_pattern(stat = "summary", fun = "mean",
                   position = position_dodge(preserve = "single"),
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,  # 0.1
                   pattern_spacing = 0.04,  # 0.025
                   pattern_key_scale_factor = 1,
                   width=0.8) + 
  scale_fill_manual(values = c("#b4a7d6ff","#f6b26bff"),labels = c("-Pi", bquote(~H[2]*O[2]))) +
  scale_pattern_manual(values = c(`0Pi` = "stripe", `2mMH2O2` = "none"),labels = c("-Pi", bquote(~H[2]*O[2]))) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.2, color = "Black", linewidth = 1, 
                position = position_dodge(width = 0.9)) +
  scale_y_continuous(breaks = seq(-0.4, 1.0, 0.2)) +
  coord_cartesian(ylim = c(-0.2, 1)) +
  geom_hline(yintercept = 1, col = "gray", linetype = "dashed") +
  xlab("promoter bashing variants") +
  ylab("Plateau") +
  guides(fill = guide_legend(nrow = 1, title = NULL), pattern = guide_legend(title = NULL)) +
  theme(strip.background = element_rect(colour="black", fill="white", size=1, linetype="solid"), 
        legend.position= "none",
        axis.title.x=element_blank())
#ggsave("output/msn4skn7_ded_plateau_0Pi.png", width = 3.3, height = 3.3)
```
